---
title: "code"
output: html_document
date: "2026-02-25"
---

```{r}
# Helper function to install CRAN packages if not already installed
install_if_missing <- function(pkg) {
  if (!requireNamespace(pkg, quietly = TRUE)) {
    install.packages(pkg, dependencies = TRUE)
  }
}

# List of CRAN packages
cran_packages <- c(
  "plyr", "doBy", "survival", "coin", "ggplot2", "lattice", "hexbin",
  "randomForest", "ggthemes", "reshape2", "ca", "FactoMineR", "MASS", "grid",
  "missMDA", "plotrix", "stats", "gdata", "psych", "corrplot", "scales",
  "RColorBrewer", "plotly", "ggsignif", "vegan", "gplots", "ggpubr",
  "dplyr", "tidyr"
)

# Install CRAN packages
lapply(cran_packages, install_if_missing)

# Install 'heatmap.2' (part of gplots, already included above)

# Install devtools if not present
install_if_missing("devtools")
library(devtools)

# Install RCM package from GitHub
if (!requireNamespace("RCM", quietly = TRUE)) {
  devtools::install_github("CenterForStatistics-UGent/RCM")
}
library(RCM)
cat("RCM package version", as.character(packageVersion("RCM")), "\n")

# Install BiocManager if needed, then use it to install phyloseq
if (!requireNamespace("BiocManager", quietly = TRUE)) {
  install.packages("BiocManager")
}
if (!requireNamespace("phyloseq", quietly = TRUE)) {
  BiocManager::install("phyloseq")
}
library(phyloseq)

# Load all packages
lapply(c(cran_packages, "RCM", "phyloseq"), library, character.only = TRUE)

```



```{r}
### Load packages
library(plyr)
library(doBy)
library(survival)
library(coin)
library(ggplot2)
library(lattice)
library(hexbin)
library(randomForest)
library(ggthemes)
library(reshape2)
library(ca)
library(FactoMineR)
library(MASS)
library(grid)
library(missMDA)
library(plotrix)
library(stats)
library(gdata)
library(psych)
library(corrplot)
library(scales)
library(RColorBrewer)
library(plotly)
library(ggsignif)
library(vegan)
library(plyr)
library(gplots)
library(devtools)
library(RCM)
library(phyloseq)
library(ggplot2)
library(ggpubr)
library(dplyr)
library(tidyr)
library(stringr)
library(readr)
library(purrr)
library(tidyverse)
library(ggcorrplot)
library(lme4)
library(lmerTest) # for p-values in lmer
library(dplyr)

```



```{r}
Shiseido_metadata_3 <- read.csv("../data/Shiseido_metadata_2.tsv", header=TRUE, sep="\t", stringsAsFactors=FALSE)
### Import exported dataframe
Shiseidodf <- read.csv("../data/Shiseidodfc.txt", header=TRUE, sep="\t", stringsAsFactors=FALSE)
alpha1 <- read_tsv("../output/alpha_diversity/217014_alpha-diversity.tsv", col_types = cols())
alpha2 <- read_tsv("../output/alpha_diversity/217015_alpha-diversity.tsv", col_types = cols())
# Step 2: Combine alpha diversity tables
alpha_combined <- full_join(alpha1, alpha2, by = "#SampleID")

# Join using SampleID_2 in metadata and #SampleID in alpha
merged_metadata_2 <- Shiseido_metadata_3 %>%
  left_join(alpha_combined, by = c("SampleID_2" = "#SampleID"))


merged_metadata_2 <- merge(
  merged_metadata_2,
  Shiseidodf[, c("X.SampleID", "Face_site")],
  by.x = "SampleID_2",
  by.y = "X.SampleID",
  all.x = TRUE  # keep all rows from merged_metadata_2
)


# Add new column with only the part before the dash
merged_metadata_2$Subject_ID <- sub("-.*", "", merged_metadata_2$microbiome_ID)


Shiseido_metadata <- merged_metadata_2
```


```{r}

fill_colors <- c(
  "Forehead" = "#ffc65e",   # warm amber
  "Nose"     = "#5693c7",   # soft blue
  "Cheek"    = "#8ccd87",   # light green
  "Ear"      = "#9dcfcf",   # muted teal (cooler, pairs with Nose & Cheek)
  "Temple"   = "#f29d1e",   # bold warm orange (near Forehead)
  "Chin"     = "#bfa6d0",   # lavender (softer contrast after orange)
  "Neck"     = "#d3d3d3"    # neutral gray (least visually dominant, good for last)
)

border_colors <- c(
  "Forehead" = "#e0a846",  # darker amber
  "Nose"     = "#3a6b99",  # darker blue
  "Cheek"    = "#4a7a4f",  # deeper green
  "Ear"      = "#6ca5a5",  # darker teal
  "Temple"   = "#c97e0f",  # deeper warm orange
  "Chin"     = "#9989ac",  # deeper lavender
  "Neck"     = "#a9a9a9"   # medium gray
)
```


# Figure 2 


```{r}
library(dplyr)
library(lme4)
library(lmerTest)
library(ggplot2)
library(ggpubr)
library(emmeans)
library(car)
library(broom)

# ── Shared themes ─────────────────────────────────────────────────────────────
shared_theme_bw <- theme_bw(base_size = 12) +
  theme(
    panel.grid       = element_blank(),
    axis.title       = element_text(size = 14),
    axis.text        = element_text(size = 12),
    plot.title       = element_text(size = 15, hjust = 0.5),
    legend.position  = "bottom",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 12),
    legend.key.size  = unit(0.8, "cm"),
    legend.spacing.x = unit(0.4, "cm")
  )

shared_theme_minimal <- theme_minimal(base_size = 12) +
  theme(
    panel.grid       = element_blank(),
    axis.line        = element_line(color = "black"),
    axis.ticks       = element_line(color = "black"),
    axis.title       = element_text(size = 14),
    axis.text        = element_text(size = 12),
    plot.title       = element_text(size = 15, hjust = 0.5),
    legend.position  = "bottom",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 12),
    legend.key.size  = unit(0.8, "cm"),
    legend.spacing.x = unit(0.4, "cm")
  )

# ── Shared helper functions ───────────────────────────────────────────────────
# Build x-axis labels with sample counts
make_x_labels <- function(df, group_var) {
  df %>%
    group_by(across(all_of(group_var))) %>%
    summarise(n = n(), .groups = "drop") %>%
    mutate(label = paste0(.data[[group_var]], "\n(n=", n, ")")) %>%
    { setNames(.$label, .[[group_var]]) }
}

make_legend_labels <- function(df, group_var) {
  df %>%
    count(.data[[group_var]]) %>%
    mutate(label = paste0(.data[[group_var]], " (n = ", n, ")")) %>%
    { setNames(.$label, .[[group_var]]) }
}
# Get pairwise label regardless of group order
get_label <- function(g1, g2, df) {
  row <- df[(df$Group1 == g1 & df$Group2 == g2) | (df$Group1 == g2 & df$Group2 == g1), ]
  if (nrow(row) == 1) row$label else NA
}

# Build annotation df for 3-group pairwise (Forehead/Nose/Cheek)
make_annotation_df <- function(wilcox_df, y_positions, y_max) {
  data.frame(
    x    = c("Forehead", "Nose",  "Forehead"),
    xend = c("Nose",     "Cheek", "Cheek"),
    label = c(
      get_label("Forehead", "Nose",  wilcox_df),
      get_label("Nose",     "Cheek", wilcox_df),
      get_label("Forehead", "Cheek", wilcox_df)
    ),
    y = y_max * y_positions
  ) %>%
    mutate(
      x_num    = match(x,    c("Forehead", "Nose", "Cheek")),
      xend_num = match(xend, c("Forehead", "Nose", "Cheek")),
      xmid     = (x_num + xend_num) / 2
    )
}

# Run Kruskal-Wallis + pairwise Wilcoxon, return tidy annotation df
run_nonparam_tests <- function(df, value_var, y_positions) {
  kruskal.test(reformulate("Face_site", value_var), data = df)

  wilcox_df <- pairwise.wilcox.test(
    df[[value_var]], df$Face_site, p.adjust.method = "BH"
  ) %>%
    broom::tidy() %>%
    rename(Group1 = group1, Group2 = group2, p_adj = p.value) %>%
    mutate(
      signif = case_when(
        p_adj < 0.001 ~ "***", p_adj < 0.01 ~ "**",
        p_adj < 0.05  ~ "*",  TRUE           ~ "ns"
      ),
      label = paste0(signif, " p = ", formatC(p_adj, format = "e", digits = 2))
    )

  make_annotation_df(wilcox_df, y_positions, max(df[[value_var]], na.rm = TRUE))
}

# ── Prepare base data ─────────────────────────────────────────────────────────
prep_avg <- function(var) {
  Shiseido_metadata %>%
    filter(Face_site %in% c("Forehead", "Nose", "Cheek")) %>%
    mutate(
      !!var := as.numeric(.data[[var]]),
      Subject_ID = as.factor(Subject_ID),
      Face_site  = factor(Face_site, levels = c("Forehead", "Nose", "Cheek"))
    ) %>%
    filter(!is.na(.data[[var]])) %>%
    group_by(Subject_ID, Face_site) %>%
    summarise(across(all_of(var), \(x) mean(x, na.rm = TRUE)), .groups = "drop")
}

# ── Plot a: Skin Redness (a*) ─────────────────────────────────────────────────
df_a    <- prep_avg("a")
x_labels_a <- make_x_labels(df_a, "Face_site")

model_aov_a <- aov(a ~ Face_site, data = df_a)
tukey_a <- TukeyHSD(model_aov_a)$Face_site %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Comparison") %>%
  mutate(
    signif = case_when(
      `p adj` < 0.001 ~ "***", `p adj` < 0.01 ~ "**",
      `p adj` < 0.05  ~ "*",  TRUE             ~ "ns"
    ),
    label = paste0(signif, " p = ", formatC(`p adj`, format = "e", digits = 2))
  )

annot_a <- data.frame(
  x    = c("Forehead", "Nose",  "Forehead"),
  xend = c("Nose",     "Cheek", "Cheek"),
  y    = c(16.0, 15.0, 15.5),
  label = tukey_a$label[match(
    c("Nose-Forehead", "Cheek-Nose", "Cheek-Forehead"), tukey_a$Comparison)]
) %>% mutate(
  x_num = match(x, c("Forehead","Nose","Cheek")),
  xend_num = match(xend, c("Forehead","Nose","Cheek")),
  xmid = (x_num + xend_num) / 2
)

Skin_Redness <- ggplot(df_a, aes(x = Face_site, y = a, fill = Face_site, color = Face_site)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 1.5),
             size = 1.5, stroke = 0.3, alpha = 0.6) +
  scale_fill_manual(values = fill_colors,    name = "Face Site") +
  scale_color_manual(values = border_colors, name = "Face Site") +
  scale_x_discrete(labels = x_labels_a) +
  labs(y = "Skin Redness (a*)", x = NULL) +
  shared_theme_bw +
  geom_segment(data = annot_a, aes(x = x, xend = xend, y = y, yend = y),
               inherit.aes = FALSE, color = "black") +
  geom_text(data = annot_a, aes(x = xmid, y = y + 0.2, label = label),
            inherit.aes = FALSE, size = 4)

# ── Plot b: Skin Brightness (L*) ──────────────────────────────────────────────
df_L    <- prep_avg("L")
x_labels_L <- make_x_labels(df_L, "Face_site")

model_aov_L <- aov(L ~ Face_site, data = df_L)
tukey_L <- TukeyHSD(model_aov_L)$Face_site %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Comparison") %>%
  mutate(
    signif = case_when(
      `p adj` < 0.001 ~ "***", `p adj` < 0.01 ~ "**",
      `p adj` < 0.05  ~ "*",  TRUE             ~ "ns"
    ),
    label = paste0(signif, " p = ", formatC(`p adj`, format = "e", digits = 2))
  )

annot_L <- data.frame(
  x    = c("Forehead", "Nose",  "Forehead"),
  xend = c("Nose",     "Cheek", "Cheek"),
  y    = c(74.0, 72.0, 73.0),
  label = tukey_L$label[match(
    c("Nose-Forehead", "Cheek-Nose", "Cheek-Forehead"), tukey_L$Comparison)]
) %>% mutate(
  x_num = match(x, c("Forehead","Nose","Cheek")),
  xend_num = match(xend, c("Forehead","Nose","Cheek")),
  xmid = (x_num + xend_num) / 2
)

Skin_Brightness <- ggplot(df_L, aes(x = Face_site, y = L, fill = Face_site, color = Face_site)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 1.5),
             size = 1.5, stroke = 0.3, alpha = 0.6) +
  scale_fill_manual(values = fill_colors,    name = "Face Site") +
  scale_color_manual(values = border_colors, name = "Face Site") +
  scale_x_discrete(labels = x_labels_L) +
  labs(y = "Skin Brightness (L*)", x = NULL) +
  shared_theme_bw +
  geom_segment(data = annot_L, aes(x = x, xend = xend, y = y, yend = y),
               inherit.aes = FALSE, color = "black") +
  geom_text(data = annot_L, aes(x = xmid, y = y + 0.4, label = label),
            inherit.aes = FALSE, size = 4)

# ── Plot c: Sebum content ─────────────────────────────────────────────────────
df_sebum    <- prep_avg("sebum")
x_labels_sebum <- make_x_labels(df_sebum, "Face_site")
annot_sebum <- run_nonparam_tests(df_sebum, "sebum", c(1.04, 0.92, 0.98))

Skin_Sebum_subset <- ggplot(df_sebum, aes(x = Face_site, y = sebum, fill = Face_site, color = Face_site)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 1.5),
             size = 1.5, stroke = 0.3, alpha = 0.6) +
  scale_fill_manual(values = fill_colors,    name = "Face Site") +
  scale_color_manual(values = border_colors, name = "Face Site") +
  scale_x_discrete(labels = x_labels_sebum) +
  labs(y = expression("Sebum content (" * mu * "g/cm"^2 * ")"), x = NULL) +
  shared_theme_bw +
  geom_segment(data = annot_sebum, aes(x = x, xend = xend, y = y, yend = y),
               inherit.aes = FALSE, color = "black") +
  geom_text(data = annot_sebum,
            aes(x = xmid, y = y + 0.025 * max(df_sebum$sebum, na.rm = TRUE), label = label),
            inherit.aes = FALSE, size = 4)

# ── Plot d: Skin Temperature ──────────────────────────────────────────────────
df_temp <- Shiseido_metadata %>%
  filter(Face_site %in% c("Forehead", "Nose", "Cheek"), !is.na(skin_temparature)) %>%
  mutate(Subject_ID = as.factor(Subject_ID),
         Face_site  = factor(Face_site, levels = c("Forehead", "Nose", "Cheek"))) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(skin_temparature = mean(skin_temparature, na.rm = TRUE), .groups = "drop")

x_labels_temp <- make_x_labels(df_temp, "Face_site")
annot_temp    <- run_nonparam_tests(df_temp, "skin_temparature", c(1.02, 1.00, 1.01))

Skin_Temp <- ggplot(df_temp, aes(x = Face_site, y = skin_temparature, fill = Face_site, color = Face_site)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 1.5),
             size = 1.5, stroke = 0.3, alpha = 0.6) +
  scale_fill_manual(values = fill_colors,    name = "Face Site") +
  scale_color_manual(values = border_colors, name = "Face Site") +
  scale_x_discrete(labels = x_labels_temp) +
  labs(y = "Skin Temperature (ºC)", x = NULL) +
  shared_theme_bw +
  geom_segment(data = annot_temp, aes(x = x, xend = xend, y = y, yend = y),
               inherit.aes = FALSE, color = "black") +
  geom_text(data = annot_temp,
            aes(x = xmid, y = y + 0.0025 * max(df_temp$skin_temparature, na.rm = TRUE), label = label),
            inherit.aes = FALSE, size = 4)

# ── Plot e: Water Content ─────────────────────────────────────────────────────
df_water <- Shiseido_metadata %>%
  filter(Face_site %in% c("Forehead", "Nose", "Cheek"), !is.na(water_content)) %>%
  mutate(Subject_ID = as.factor(Subject_ID),
         Face_site  = factor(Face_site, levels = c("Forehead", "Nose", "Cheek"))) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(water_content = mean(water_content, na.rm = TRUE), .groups = "drop")

x_labels_water <- make_x_labels(df_water, "Face_site")
annot_water    <- run_nonparam_tests(df_water, "water_content", c(1.10, 1.00, 1.05))

Skin_WaterContent <- ggplot(df_water, aes(x = Face_site, y = water_content, fill = Face_site, color = Face_site)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 1.5),
             size = 1.5, stroke = 0.3, alpha = 0.6) +
  scale_fill_manual(values = fill_colors,    name = "Face Site") +
  scale_color_manual(values = border_colors, name = "Face Site") +
  scale_x_discrete(labels = x_labels_water) +
  labs(y = "Water Content (a.u.)", x = NULL) +
  shared_theme_bw +
  geom_segment(data = annot_water, aes(x = x, xend = xend, y = y, yend = y),
               inherit.aes = FALSE, color = "black") +
  geom_text(data = annot_water,
            aes(x = xmid, y = y + 0.03 * max(df_water$water_content, na.rm = TRUE), label = label),
            inherit.aes = FALSE, size = 4)

# ── Helper: LMM scatter plot ──────────────────────────────────────────────────
make_lmm_scatter <- function(df, x_var, y_var, x_lab, y_lab, title) {
  model  <- lmer(reformulate(c(x_var, "Face_site", "(1 | Subject_ID)"), y_var), data = df)
  smry   <- summary(model)$coefficients
  cor_text <- paste0("β = ", round(smry[x_var, "Estimate"], 2),
                     ", p = ", signif(smry[x_var, "Pr(>|t|)"], 2))
  leg_labels <- make_legend_labels(df, "Face_site")

  ggplot(df, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(aes(fill = Face_site, color = Face_site),
               shape = 21, size = 3, alpha = 0.6, stroke = 0.6) +
    geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.7) +
    annotate("text", x = Inf, y = Inf, label = cor_text,
             hjust = 1.1, vjust = 1.5, size = 4) +
    scale_fill_manual(values = fill_colors,    labels = leg_labels, name = "Face Site") +
    scale_color_manual(values = border_colors, labels = leg_labels, name = "Face Site") +
    labs(x = x_lab, y = y_lab, title = title) +
    shared_theme_minimal
}

# ── Plots f–h: LMM scatter plots ─────────────────────────────────────────────
df_La <- Shiseido_metadata %>%
  mutate(L = as.numeric(L), a = as.numeric(a),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(L), !is.na(a)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_L = mean(L, na.rm = TRUE), mean_a = mean(a, na.rm = TRUE), .groups = "drop")

df_sebum_a <- Shiseido_metadata %>%
  mutate(sebum = as.numeric(sebum), a = as.numeric(a),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(sebum), !is.na(a)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_sebum = mean(sebum, na.rm = TRUE), mean_a = mean(a, na.rm = TRUE), .groups = "drop")

df_sebum_L <- Shiseido_metadata %>%
  mutate(sebum = as.numeric(sebum), L = as.numeric(L),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(sebum), !is.na(L)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_sebum = mean(sebum, na.rm = TRUE), mean_L = mean(L, na.rm = TRUE), .groups = "drop")

e_LM_avg <- make_lmm_scatter(df_La, "mean_L", "mean_a",
  "Skin Brightness (L*)", "Skin Redness (a*)",
  "Association between skin brightness and redness")

skinredness_sebum_LM_avg <- make_lmm_scatter(df_sebum_a, "mean_sebum", "mean_a",
  "Sebum content (µg/cm²)", "Skin Redness (a*)",
  "Association between sebum content and skin redness")

skinbrightness_sebum_LM_avg <- make_lmm_scatter(df_sebum_L, "mean_sebum", "mean_L",
  "Sebum content (µg/cm²)", "Skin Brightness (L*)",
  "Association between sebum content and skin brightness")

# ── Plot i: Sebum by Subject (Kruskal-Wallis) ─────────────────────────────────
metadata_with_runorder <- read.csv("../data/metadata_with_runorder_050132025.csv")

q           <- quantile(metadata_with_runorder$sebum, probs = c(0.25, 0.75), na.rm = TRUE)
low_thresh  <- q[1]; high_thresh <- q[2]

df_avg_sebum <- metadata_with_runorder %>%
  filter(!is.na(Subject_ID), !is.na(sebum),
         Face_site %in% c("Forehead", "Nose", "Cheek")) %>%
  mutate(Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(sebum = mean(sebum, na.rm = TRUE), .groups = "drop")

subject_median_info <- df_avg_sebum %>%
  group_by(Subject_ID) %>%
  summarise(median_sebum = median(sebum, na.rm = TRUE)) %>%
  mutate(sebum_group_median = case_when(
    median_sebum < low_thresh  ~ "low",
    median_sebum > high_thresh ~ "high",
    TRUE                       ~ "medium"
  ))

df_avg_sebum <- df_avg_sebum %>%
  left_join(subject_median_info, by = "Subject_ID") %>%
  mutate(
    sebum_group_median = factor(sebum_group_median, levels = c("low", "medium", "high")),
    Subject_ID = factor(Subject_ID,
      levels = subject_median_info %>% arrange(median_sebum) %>% pull(Subject_ID))
  )

kw_p <- kruskal.test(sebum ~ Subject_ID, data = df_avg_sebum)$p.value

group_counts  <- df_avg_sebum %>%
  group_by(sebum_group_median) %>% summarise(n = n()) %>%
  mutate(label = case_when(
    sebum_group_median == "low"    ~ paste0("Low (< ", low_thresh,  " µg/cm²) (n=", n, ")"),
    sebum_group_median == "medium" ~ paste0("Medium (", low_thresh, "–", high_thresh, " µg/cm²) (n=", n, ")"),
    sebum_group_median == "high"   ~ paste0("High (> ", high_thresh," µg/cm²) (n=", n, ")")
  ))
legend_labels <- setNames(group_counts$label, group_counts$sebum_group_median)

Skin_Sebum_KW_subjects <- ggplot(df_avg_sebum,
    aes(y = sebum, x = Subject_ID, fill = sebum_group_median, color = sebum_group_median)) +
  geom_boxplot(outlier.shape = NA, color = "black") +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.5),
             size = 1.5, stroke = 0.3, alpha = 0.6, color = "black") +
  scale_fill_manual(values  = c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214"),
                    name    = "Sebum Group (Median)", labels = legend_labels) +
  scale_color_manual(values = c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214"),
                     guide  = "none") +
  scale_y_log10() +
  labs(y = expression("Sebum content (" * mu * "g/cm"^2 * ")"), x = "Subject ID") +
  shared_theme_bw +
  theme(axis.text.x = element_text(size = 10, angle = 45, hjust = 1)) +
  annotate("text", x = 7, y = Inf,
           label = paste0("Kruskal-Wallis, p = ", format(kw_p, digits = 3, scientific = TRUE)),
           vjust = 1.5, size = 4)

# ── Assemble Figure 2 ─────────────────────────────────────────────────────────
boxplot_fg <- ggarrange(
  Skin_Redness, Skin_Brightness, Skin_Sebum_subset, Skin_Temp, Skin_WaterContent,
  ncol = 5, labels = c("a","b","c","d","e"),
  common.legend = TRUE, legend = "bottom",
  font.label = list(size = 18, face = "bold")
)

corrplot <- ggarrange(
  e_LM_avg, skinredness_sebum_LM_avg, skinbrightness_sebum_LM_avg,
  ncol = 3, labels = c("f","g","h"),
  common.legend = TRUE, legend = "bottom",
  font.label = list(size = 18, face = "bold")
)

top_grid <- ggarrange(
  boxplot_fg, corrplot, Skin_Sebum_KW_subjects,
  ncol = 1, nrow = 3,
  labels = c("", "", "i"),
  heights = c(1, 1, 0.75),
  font.label = list(size = 18, face = "bold")
)


```

```{r}
# ── Save ──────────────────────────────────────────────────────────────────────
ggsave("../figures/main/Figure2.png", plot = top_grid, width = 17, height = 14, dpi = 300)
ggsave("../figures/main/Figure2.svg", plot = top_grid, width = 17, height = 14)
```

# Figure 3


```{r}
# ── Shared theme ──────────────────────────────────────────────────────────────
shared_theme <- theme_minimal(base_size = 12) +
  theme(
    panel.grid       = element_blank(),
    axis.line        = element_line(color = "black"),
    axis.ticks       = element_line(color = "black"),
    axis.title       = element_text(size = 14),
    axis.text        = element_text(size = 12),
    plot.title       = element_text(size = 15, hjust = 0.5),
    legend.position  = "bottom",
    legend.title     = element_text(size = 12),
    legend.text      = element_text(size = 12),
    legend.key.size  = unit(0.8, "cm"),
    legend.spacing.x = unit(0.4, "cm")
  )

# ── Prepare data ──────────────────────────────────────────────────────────────
df_lmm <- Shiseido_metadata %>%
  filter(Face_site %in% c("Forehead", "Nose", "Cheek"),
         !is.na(shannon_entropy),
         Subject_ID != "na") %>%
  mutate(
    Subject_ID = as.factor(Subject_ID),
    Face_site  = factor(Face_site, levels = c("Forehead", "Nose", "Cheek"))
  )

# ── x-axis labels with counts (for boxplot) ───────────────────────────────────
x_labels <- df_lmm %>%
  group_by(Face_site) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(label = paste0(Face_site, "\n(n=", n, ")")) %>%
  with(setNames(label, Face_site))

# ── Legend labels with counts (for scatter plot) ──────────────────────────────
site_counts      <- table(df_lmm$Face_site)
face_site_labels <- setNames(
  paste0(names(site_counts), " (n=", site_counts, ")"),
  names(site_counts)
)

# ── LMM: Shannon entropy ~ Face site ─────────────────────────────────────────
model_lmm <- lmer(shannon_entropy ~ Face_site + (1 | Subject_ID), data = df_lmm)
anova(model_lmm)

pairwise_df <- pairs(emmeans(model_lmm, ~ Face_site), adjust = "BH") %>%
  as.data.frame() %>%
  mutate(
    Group1  = sub(" - .*", "", contrast),
    Group2  = sub(".*- ",  "", contrast),
    signif  = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE            ~ "ns"
    ),
    label = paste0(signif, " p = ", formatC(p.value, format = "e", digits = 2)),
    x     = match(Group1, levels(df_lmm$Face_site)),
    xend  = match(Group2, levels(df_lmm$Face_site)),
    xmid  = (x + xend) / 2
  )

y_max          <- max(df_lmm$shannon_entropy, na.rm = TRUE)
pairwise_df$y  <- y_max * seq(1.10, 1.00, length.out = nrow(pairwise_df))

# ── Plot A: Shannon entropy boxplot ───────────────────────────────────────────
Shannon_entropy_plot_lmm <- ggplot(df_lmm,
    aes(x = Face_site, y = shannon_entropy, fill = Face_site, color = Face_site)) +
  geom_boxplot(outlier.shape = NA) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 1.5),
             size = 1.5, stroke = 0.3, alpha = 0.6) +
  scale_fill_manual(values = fill_colors,   labels = face_site_labels) +
  scale_color_manual(values = border_colors, labels = face_site_labels) +
  scale_x_discrete(labels = x_labels) +
  labs(y = "Shannon Entropy", x = NULL,
       title = "16S rRNA (V4) Alpha Diversity",
       fill = "Face Site", color = "Face Site") +
  shared_theme +
  geom_segment(data = pairwise_df,
               aes(x = x, xend = xend, y = y, yend = y),
               inherit.aes = FALSE, color = "black") +
  geom_text(data = pairwise_df,
            aes(x = xmid, y = y + 0.02 * y_max, label = label),
            inherit.aes = FALSE, size = 4)

# ── LMM: Sebum ~ Shannon + Face site ─────────────────────────────────────────
model_lmm_sebum <- lmer(sebum ~ shannon_entropy + Face_site + (1 | Subject_ID),
                        data = df_lmm)
summary(model_lmm_sebum)

sebum_coef <- summary(model_lmm_sebum)$coefficients["shannon_entropy", ]
cor_text   <- paste0("β = ", round(sebum_coef["Estimate"], 2),
                     ", p = ", signif(sebum_coef["Pr(>|t|)"], 2))

# ── Plot B: Sebum ~ Shannon scatter ───────────────────────────────────────────
sebum_shannon_entropy_lmm <- ggplot(df_lmm, aes(x = shannon_entropy, y = sebum)) +
  geom_point(aes(fill = Face_site, color = Face_site),
             shape = 21, size = 3, alpha = 0.6, stroke = 0.6) +
  geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.7) +
  annotate("text", x = Inf, y = Inf, label = cor_text,
           hjust = 1.1, vjust = 1.5, size = 4) +
  scale_fill_manual(values = fill_colors,    labels = face_site_labels) +
  scale_color_manual(values = border_colors, labels = face_site_labels) +
  labs(x = "Shannon Entropy",
       y = expression("Sebum content (" * mu * "g/cm"^2 * ")"),
       title = "Association between Shannon Entropy and Sebum Content",
       fill = "Face Site", color = "Face Site") +
  shared_theme


# ── Assemble Figure 3 ─────────────────────────────────────────────────────────
boxplot_fg <- ggarrange(
  Shannon_entropy_plot_lmm, sebum_shannon_entropy_lmm,
  ncol = 2, nrow = 1,
  labels = c("a", "b"),
  common.legend = FALSE,
  font.label = list(size = 18, face = "bold")
)

empty_fg <- ggarrange(
  NULL, NULL,
  ncol = 2, nrow = 1,
  labels = c("c", "d"),
  font.label = list(size = 18, face = "bold")
)

top_grid <- ggarrange(
  boxplot_fg, empty_fg,
  ncol = 1, nrow = 2,
  heights = c(1, 1),
  font.label = list(size = 18, face = "bold")
)
```


```{r}
# ── Save ──────────────────────────────────────────────────────────────────────
ggsave("../figures/main/Figure3.png", plot = top_grid,
       width = 13, height = 11, dpi = 300)
ggsave("../figures/main/Figure3.svg", plot = top_grid,
       width = 13, height = 11)
```


# Supplementary Figure 1

```{r}
# ── Helper: age correlation scatter ──────────────────────────────────────────
make_age_cor_scatter <- function(df, x_var, y_var, y_lab, title) {
  p_x    <- shapiro.test(df[[x_var]])$p.value
  p_y    <- shapiro.test(df[[y_var]])$p.value
  method <- ifelse(p_x > 0.05 & p_y > 0.05, "pearson", "spearman")

  cor_test  <- cor.test(df[[x_var]], df[[y_var]], method = method)
  cor_label <- ifelse(method == "pearson", "Pearson r", "Spearman ρ")
  cor_text  <- paste0(cor_label, " = ", round(cor_test$estimate, 2),
                      ", p = ", signif(cor_test$p.value, 2))

  leg_labels <- make_legend_labels(df, "Face_site")

  ggplot(df, aes(x = .data[[x_var]], y = .data[[y_var]])) +
    geom_point(aes(fill = Face_site, color = Face_site),
               shape = 21, size = 3, alpha = 0.6, stroke = 0.6) +
    geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.7) +
    annotate("text", x = Inf, y = Inf, label = cor_text,
             hjust = 1.1, vjust = 1.5, size = 4) +
    scale_fill_manual(values = fill_colors,    labels = leg_labels, name = "Face Site") +
    scale_color_manual(values = border_colors, labels = leg_labels, name = "Face Site") +
    labs(x = "Age (years)", y = y_lab, title = title) +
    shared_theme_minimal
}

# ── Prepare data ──────────────────────────────────────────────────────────────
df_temp_a <- Shiseido_metadata %>%
  mutate(skin_temparature = as.numeric(skin_temparature), a = as.numeric(a),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(skin_temparature), !is.na(a)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_temp = mean(skin_temparature, na.rm = TRUE),
            mean_a    = mean(a,                na.rm = TRUE), .groups = "drop")

df_temp_L <- Shiseido_metadata %>%
  mutate(skin_temparature = as.numeric(skin_temparature), L = as.numeric(L),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(skin_temparature), !is.na(L)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_temp = mean(skin_temparature, na.rm = TRUE),
            mean_L    = mean(L,                na.rm = TRUE), .groups = "drop")

df_water_a <- Shiseido_metadata %>%
  mutate(water_content = as.numeric(water_content), a = as.numeric(a),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(water_content), !is.na(a)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_water = mean(water_content, na.rm = TRUE),
            mean_a     = mean(a,             na.rm = TRUE), .groups = "drop")

df_water_L <- Shiseido_metadata %>%
  mutate(water_content = as.numeric(water_content), L = as.numeric(L),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(water_content), !is.na(L)) %>%
  group_by(Subject_ID, Face_site) %>%
  summarise(mean_water = mean(water_content, na.rm = TRUE),
            mean_L     = mean(L,             na.rm = TRUE), .groups = "drop")

df_R2 <- Shiseido_metadata %>%
  mutate(age = as.numeric(age), cutemeter_R2 = as.numeric(cutemeter_R2),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(age), !is.na(cutemeter_R2)) %>%
  group_by(Subject_ID, age, Face_site) %>%
  summarise(mean_R2 = mean(cutemeter_R2, na.rm = TRUE), .groups = "drop")

df_R7 <- Shiseido_metadata %>%
  mutate(age = as.numeric(age), cutemeter_R7 = as.numeric(cutemeter_R7),
         Subject_ID = as.factor(Subject_ID), Face_site = as.factor(Face_site)) %>%
  filter(!is.na(age), !is.na(cutemeter_R7)) %>%
  group_by(Subject_ID, age, Face_site) %>%
  summarise(mean_R7 = mean(cutemeter_R7, na.rm = TRUE), .groups = "drop")

# ── Plots a–d: LMM scatter ────────────────────────────────────────────────────
d_LM_avg <- make_lmm_scatter(df_temp_a, "mean_temp", "mean_a",
  "Skin Temperature (°C)", "Skin Redness (a*)",
  "Association between average skin temperature and skin redness")

c_LM_avg <- make_lmm_scatter(df_temp_L, "mean_temp", "mean_L",
  "Skin Temperature (°C)", "Skin Brightness (L*)",
  "Association between average skin temperature and skin brightness")

skinredness_watercontent_LM_avg <- make_lmm_scatter(df_water_a, "mean_water", "mean_a",
  "Water Content (a.u.)", "Skin Redness (a*)",
  "Association between water content and skin redness")

skinbrightness_watercontent_LM_avg <- make_lmm_scatter(df_water_L, "mean_water", "mean_L",
  "Water Content (a.u.)", "Skin Brightness (L*)",
  "Association between water content and skin brightness")

# ── Plots e–f: Age correlation ────────────────────────────────────────────────
r2_plot <- make_age_cor_scatter(df_R2, "age", "mean_R2",
  "Cutometer R2 (elasticity index)",
  "Correlation between age and cutometer R2")

r7_plot <- make_age_cor_scatter(df_R7, "age", "mean_R7",
  "Cutometer R7 (elastic recovery index)",
  "Correlation between age and cutometer R7")


top_grid_suppl1 <- ggarrange(
  d_LM_avg,
  c_LM_avg,
  skinredness_watercontent_LM_avg,
  skinbrightness_watercontent_LM_avg,
  r2_plot,
  r7_plot,
  ncol          = 2,
  nrow          = 3,
  labels        = c("a","b","c","d","e","f"),
  common.legend = TRUE,
  legend        = "bottom",
  font.label    = list(size = 18, face = "bold")
)


```


```{r}
# PNG
png("../figures/supplementary/suppl_fig1.png", width = 14, height = 13, 
    units = "in", res = 300)
print(top_grid_suppl1)
dev.off()

# SVG
svg("../figures/supplementary/suppl_fig1.svg", width = 14, height = 13)
print(top_grid_suppl1)
dev.off()
```

Supplementary Figure 4-9

```{r}
# Required libraries
library(ggplot2)
library(dplyr)
library(tidyr)
library(rstatix)
library(ggpubr)
library(lme4)
library(broom.mixed)
library(performance)
library(patchwork)

# ── Define x variables and labels ────────────────────────────────────────────
x_vars <- c("water_content", "sebum", "a", "L", "skin_temparature")

x_axis_labels <- c(
  "L"                = "Skin Brightness (L*)",
  "a"                = "Skin Redness (a*)",
  "b"                = "Skin Yellowness (b*)",
  "sebum"            = "Sebum concentration (µg/cm²)",
  "water_content"    = "Water Content (a.u.)",
  "skin_temparature" = "Skin Temperature (°C)",
  "age"              = "Age (years)"
)

# ── Define amino acids ────────────────────────────────────────────────────────
amino_acids <- c("Asp", "Glu", "hyP", "Ser", "Asn", "Gly", "Gln", "Thr", "Ala", "Tau",
                 "His", "Cit", "Pro", "Arg", "Val", "Tyr", "Met", "Cys2", "Ile", "Leu",
                 "Orn", "Lys", "Phe", "Trp", "Total_AAs")
amino_acids_wo_protein <- paste0(amino_acids, "_wo_Protein")
Shiseido_metadata$Total_AAs_wo_Protein <- Shiseido_metadata$Total_Aas_wo_Protein

# ── y-axis labels ─────────────────────────────────────────────────────────────
y_axis_labels_wo_protein <- c(
  "Asp_wo_Protein"            = "Aspartic Acid",
  "Glu_wo_Protein"            = "Glutamic Acid",
  "hyP_wo_Protein"            = "Hydroxyproline",
  "Ser_wo_Protein"            = "Serine",
  "Asn_wo_Protein"            = "Asparagine",
  "Gly_wo_Protein"            = "Glycine",
  "Gln_wo_Protein"            = "Glutamine",
  "Thr_wo_Protein"            = "Threonine",
  "Ala_wo_Protein"            = "Alanine",
  "Tau_wo_Protein"            = "Taurine",
  "His_wo_Protein"            = "Histidine",
  "Cit_wo_Protein"            = "Citrulline",
  "Pro_wo_Protein"            = "Proline",
  "Arg_wo_Protein"            = "Arginine",
  "Val_wo_Protein"            = "Valine",
  "Tyr_wo_Protein"            = "Tyrosine",
  "Met_wo_Protein"            = "Methionine",
  "Cys2_wo_Protein"           = "Cystine",
  "Ile_wo_Protein"            = "Isoleucine",
  "Leu_wo_Protein"            = "Leucine",
  "Orn_wo_Protein"            = "Ornithine",
  "Lys_wo_Protein"            = "Lysine",
  "Phe_wo_Protein"            = "Phenylalanine",
  "Trp_wo_Protein"            = "Tryptophan",
  "Total_AAs_wo_Protein"      = "Total Amino Acids",
  "Urea_wo_Protein"           = "Urea",
  "Creatinine_wo_Protein"     = "Creatinine",
  "UCA_wo_Protein"            = "Urocanic Acid",
  "LA_wo_Protein"             = "Lactic Acid",
  "PCA_wo_Protein"            = "Pyrrolidone Carboxylic Acid",
  "D.glucosamine_wo_Protein"  = "D-Glucosamine",
  "Total_org._NMF_wo_Protein" = "Total Organic NMF"
)

# ── Model comparison (AIC: raw vs log-transformed) ────────────────────────────
model_comparison_results <- list()
valid_face_sites <- c("Forehead", "Nose", "Cheek")

for (x in x_vars) {
  for (y in amino_acids_wo_protein) {
    sub_df <- Shiseido_metadata %>%
      filter(Face_site %in% valid_face_sites,
             !is.na(Subject_ID), Subject_ID != "na") %>%
      select(Subject_ID, Face_site, all_of(c(x, y))) %>%
      filter(!is.na(.data[[x]]), !is.na(.data[[y]])) %>%
      mutate(Subject_ID = as.factor(Subject_ID),
             Face_site  = as.factor(Face_site))

    if (nrow(sub_df) < 5) next

    model_raw <- tryCatch(
      lmer(as.formula(paste(y, "~", x, "+ Face_site + (1|Subject_ID)")), data = sub_df),
      error = function(e) NULL
    )

    sub_df$log_x <- log10(pmax(as.numeric(sub_df[[x]]), 1))
    sub_df$log_y <- log10(pmax(as.numeric(sub_df[[y]]), 1))

    model_log <- tryCatch(
      lmer(log_y ~ log_x + Face_site + (1|Subject_ID), data = sub_df),
      error = function(e) NULL
    )

    if (!is.null(model_raw) && !is.null(model_log)) {
      aic_raw   <- AIC(model_raw)
      aic_log   <- AIC(model_log)
      delta_aic <- aic_raw - aic_log

      model_comparison_results[[paste(x, y, sep = "_")]] <- data.frame(
        X            = x,
        Y            = y,
        AIC_raw      = aic_raw,
        AIC_log      = aic_log,
        Delta_AIC    = delta_aic,
        Better_Model = ifelse(delta_aic > 2, "Log-transformed", "Untransformed or Similar")
      )
    }
  }
}

model_comparison_df <- do.call(rbind, model_comparison_results)

# ── Plotting function ─────────────────────────────────────────────────────────
output_dir <- "../figures/supplementary/lmm"
dir.create(output_dir, showWarnings = FALSE)

plot_lmm_by_heatmap_model <- function(x_var, model_comparison_df, metadata_df) {
  message("Processing: ", x_var)

  plot_list     <- list()
  relevant_rows <- model_comparison_df %>% filter(X == x_var)

  for (i in seq_len(nrow(relevant_rows))) {
    y_var        <- relevant_rows$Y[i]
    better_model <- relevant_rows$Better_Model[i]

    sub_df <- metadata_df %>%
      filter(Face_site %in% c("Forehead", "Nose", "Cheek"),
             !is.na(Subject_ID), Subject_ID != "na") %>%
      select(Subject_ID, Face_site, all_of(c(x_var, y_var))) %>%
      filter(!is.na(.data[[x_var]]), !is.na(.data[[y_var]])) %>%
      mutate(Subject_ID = as.factor(Subject_ID),
             Face_site  = as.factor(Face_site))

    if (nrow(sub_df) < 5) next

    # Legend labels with counts
    legend_info <- sub_df %>%
      group_by(Face_site) %>%
      summarise(n = n(), .groups = "drop") %>%
      mutate(label = paste0(Face_site, " (n = ", n, ")"))
    sub_df$Face_site <- factor(sub_df$Face_site, levels = legend_info$Face_site)
    legend_labels    <- setNames(legend_info$label, legend_info$Face_site)

    # ── Build model and base plot ─────────────────────────────────────────────
    if (better_model == "Log-transformed") {
      sub_df$log_x <- log10(pmax(as.numeric(sub_df[[x_var]]), 1))
      sub_df$log_y <- log10(pmax(as.numeric(sub_df[[y_var]]), 1))

      model <- tryCatch(
        lmer(log_y ~ log_x + Face_site + (1 | Subject_ID), data = sub_df),
        error = function(e) NULL
      )
      term      <- "log_x"
      x_label <- bquote(log[10](.(x_axis_labels[[x_var]])))
      y_label <- bquote(log[10](.(y_axis_labels_wo_protein[[y_var]]) ~ "[pmol/cm²]"))
      base_plot <- ggplot(sub_df, aes(x = log_x, y = log_y))

    } else {
      model <- tryCatch(
        lmer(as.formula(paste(y_var, "~", x_var, "+ Face_site + (1 | Subject_ID)")),
             data = sub_df),
        error = function(e) NULL
      )
      term      <- x_var

      x_label <- bquote(log[10](.(x_axis_labels[[x_var]])))
      y_label <- bquote(log[10](.(y_axis_labels_wo_protein[[y_var]]) ~ "[pmol/cm²]"))

      base_plot <- ggplot(sub_df, aes(x = .data[[x_var]], y = .data[[y_var]]))
    }

    if (!is.null(model)) {
      model_summary <- broom.mixed::tidy(model)
      effect_row    <- model_summary %>% filter(term == !!term)

      if (nrow(effect_row) == 1) {
        cor_text <- paste0("β = ", round(effect_row$estimate, 2),
                           ", p = ", signif(effect_row$p.value, 2))

        p <- base_plot +
          geom_point(aes(fill = Face_site, color = Face_site),
                     shape = 21, size = 2.5, alpha = 0.6, stroke = 0.5) +
          geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.6) +
          annotate("text", x = Inf, y = Inf, label = cor_text,
                   hjust = 1.1, vjust = 1.5, size = 3.5) +
          scale_fill_manual(values = fill_colors,    name = "Face Site", labels = legend_labels) +
          scale_color_manual(values = border_colors, guide = "none") +
          labs(title = y_axis_labels_wo_protein[[y_var]], x = x_label, y = y_label) +
          theme_minimal(base_size = 12) +
          theme(
            panel.grid = element_blank(),
            axis.line  = element_line(color = "black"),
            axis.ticks = element_line(color = "black"),
            axis.text  = element_text(size = 10),
            axis.title = element_text(size = 11),
            plot.title = element_text(size = 12, hjust = 0.5)
          )

        plot_list[[length(plot_list) + 1]] <- p
      }
    }
  }

  if (length(plot_list) > 0) {
    # Apply legend theme to each plot individually — avoids & operator issues
    plot_list <- lapply(plot_list, function(p) {
      p + theme(
        legend.position  = "bottom",
        legend.direction = "horizontal",
        legend.title     = element_text(size = 11),
        legend.text      = element_text(size = 10),
        plot.tag         = element_text(size = 14, face = "bold")
      )
    })

    combined_plot <- wrap_plots(plot_list, ncol = 4, guides = "collect") +
      plot_annotation(
        title      = paste0("Amino Acid / NMF vs ", x_axis_labels[[x_var]],
                            " (LMM adjusted for Face Site and Subject ID)"),
        tag_levels = "a",
        theme      = theme(                        # ← controls the overall layout theme
          legend.position  = "bottom",
          legend.direction = "horizontal",
          plot.title       = element_text(size = 14, hjust = 0.5)
        )
      )


    png(file.path(output_dir, paste0(x_var, "_vs_aminoacids.png")),
        width = 16, height = 20, units = "in", res = 300)
    print(combined_plot)
    dev.off()

    svg(file.path(output_dir, paste0(x_var, "_vs_aminoacids.svg")),
        width = 16, height = 20)
    print(combined_plot)
    dev.off()
  }
}

# ── Run for all predictors ────────────────────────────────────────────────────
for (x in x_vars) {
  plot_lmm_by_heatmap_model(
    x_var               = x,
    model_comparison_df = model_comparison_df,
    metadata_df         = Shiseido_metadata
  )
}
```

#Supplementary Figure 17-18

```{r}
# Adapt for those 

# Define column names
amino_acids <- c("Urea","Creatinine","UCA","LA","PCA")#,"Total_org._NMF")

amino_acids_wo_protein <- paste0(amino_acids, "_wo_Protein")
```


```{r}
# ── Amino acid definitions ────────────────────────────────────────────────────
amino_acids <- c("Asp", "Glu", "hyP", "Ser", "Asn", "Gly", "Gln", "Thr", "Ala", "Tau",
                 "His", "Cit", "Pro", "Arg", "Val", "Tyr", "Met", "Cys2", "Ile", "Leu",
                 "Orn", "Lys", "Phe", "Trp", "Total_AAs")
amino_acids_wo_protein <- paste0(amino_acids, "_wo_Protein")
Shiseido_metadata$Total_AAs_wo_Protein <- Shiseido_metadata$Total_Aas_wo_Protein
```



```{r}
# ── Plotting function ─────────────────────────────────────────────────────────
plot_amino_entropy <- function(df, col_name, label, normalized = TRUE) {
  df <- df %>%
    mutate(
      x_raw = as.numeric(shannon_entropy),
      y_raw = as.numeric(.data[[col_name]]),
      log_x = log10(x_raw + 0.01),
      log_y = log10(y_raw + 0.01)
    ) %>%
    filter(!is.na(log_x), !is.na(log_y))

  if (nrow(df) < 3 || length(unique(df$log_y)) <= 1) {
    message(paste("Skipping", col_name, "- not enough valid data or no variability."))
    return(NULL)
  }

  # Normality tests on untransformed data to choose method
  p_x <- tryCatch(shapiro.test(df$x_raw)$p.value, error = function(e) NA)
  p_y <- tryCatch(shapiro.test(df$y_raw)$p.value, error = function(e) NA)
  method <- if (!is.na(p_x) && !is.na(p_y) && p_x > 0.05 && p_y > 0.05) "pearson" else "spearman"

  cor_result <- tryCatch(
    cor.test(df$log_x, df$log_y, method = method),
    error = function(e) NULL
  )

  if (is.null(cor_result)) {
    message(paste("Skipping", col_name, "- correlation test failed."))
    return(NULL)
  }

  cor_label <- if (method == "pearson") "Pearson r" else "Spearman \u03c1"
  cor_text  <- paste0(cor_label, " = ", round(cor_result$estimate, 2),
                      ", p = ", signif(cor_result$p.value, 2))

  y_unit  <- if (normalized) "nmol/mg protein" else "pmol/cm\u00b2"
  y_label <- bquote(log[10]*"("*.(label)*" ["*.(y_unit)*"])")

  ggplot(df, aes(x = log_x, y = log_y)) +
    geom_point(aes(fill = Face_site, color = Face_site),
               shape = 21, size = 3, alpha = 0.6, stroke = 0.6) +
    geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.7) +
    annotate("text", x = Inf, y = Inf, label = cor_text,
             hjust = 1.1, vjust = 1.5, size = 4) +
    scale_fill_manual(values = fill_colors,    name = "Face Site") +
    scale_color_manual(values = border_colors, name = "Face Site") +
    labs(
      x     = bquote(log[10]*"(Shannon Entropy)"),
      y     = y_label,
      title = paste("Correlation with", label)
    ) +
    theme_minimal(base_size = 14) +
    theme(
      panel.grid   = element_blank(),
      axis.line    = element_line(color = "black"),
      axis.ticks   = element_line(color = "black"),
      axis.title   = element_text(size = 18),
      axis.text    = element_text(size = 16),
      plot.title   = element_text(size = 20, hjust = 0.5),
      legend.position  = "bottom",
      legend.title     = element_text(size = 17),
      legend.text      = element_text(size = 16),
      legend.key.size  = unit(0.8, "cm"),
      legend.spacing.x = unit(0.4, "cm")
    )
}

# ── Build plot lists ──────────────────────────────────────────────────────────
plots <- Filter(Negate(is.null), lapply(amino_acids, function(col) {
  plot_amino_entropy(Shiseido_metadata, col, col, normalized = TRUE)
}))

plots_wo <- Filter(Negate(is.null), lapply(amino_acids_wo_protein, function(col) {
  label <- gsub("_wo_Protein", "", col)
  plot_amino_entropy(Shiseido_metadata, col, label, normalized = FALSE)
}))

# ── Assemble — apply theme per plot to avoid & precedence issue ───────────────
apply_tag_theme <- function(plot_list) {
  lapply(plot_list, function(p) {
    p + theme(plot.tag = element_text(size = 20, face = "bold"))
  })
}



```


```{r}

combined_plot <- wrap_plots(apply_tag_theme(plots), ncol = 4, guides = "collect") +
  plot_annotation(
    tag_levels = "a",
    theme = theme(
      legend.position  = "bottom",
      legend.direction = "horizontal",
      legend.title     = element_text(size = 17),
      legend.text      = element_text(size = 16),
      legend.key.size  = unit(0.8, "cm"),
      legend.spacing.x = unit(0.4, "cm")
    )
  )

combined_plot_wo <- wrap_plots(apply_tag_theme(plots_wo), ncol = 4, guides = "collect") +
  plot_annotation(
    tag_levels = "a",
    theme = theme(
      legend.position  = "bottom",
      legend.direction = "horizontal",
      legend.title     = element_text(size = 17),
      legend.text      = element_text(size = 16),
      legend.key.size  = unit(0.8, "cm"),
      legend.spacing.x = unit(0.4, "cm")
    )
  )
# ── Save ──────────────────────────────────────────────────────────────────────
dir.create("../figures/supplementary/spearman", showWarnings = FALSE, recursive = TRUE)

png("../figures/supplementary/spearman/Shannon_vs_AminoAcids_log.png",
    width = 20, height = 28, units = "in", res = 300)
print(combined_plot)
dev.off()

png("../figures/supplementary/spearman/Shannon_vs_AminoAcids_no_protein_log.png",
    width = 22, height = 28, units = "in", res = 300)
print(combined_plot_wo)
dev.off()

svg("../figures/supplementary/spearman/Shannon_vs_AminoAcids_log.svg",
    width = 20, height = 28)
print(combined_plot)
dev.off()

svg("../figures/supplementary/spearman/Shannon_vs_AminoAcids_no_protein_log.svg",
    width = 22, height = 28)
print(combined_plot_wo)
dev.off()
```

```{r}

combined_plot <- wrap_plots(apply_tag_theme(plots), ncol = 2, guides = "collect") +
  plot_annotation(
    tag_levels = "a",
    theme = theme(
      legend.position  = "bottom",
      legend.direction = "horizontal",
      legend.title     = element_text(size = 17),
      legend.text      = element_text(size = 16),
      legend.key.size  = unit(0.8, "cm"),
      legend.spacing.x = unit(0.4, "cm")
    )
  )

combined_plot_wo <- wrap_plots(apply_tag_theme(plots_wo), ncol = 2, guides = "collect") +
  plot_annotation(
    tag_levels = "a",
    theme = theme(
      legend.position  = "bottom",
      legend.direction = "horizontal",
      legend.title     = element_text(size = 17),
      legend.text      = element_text(size = 16),
      legend.key.size  = unit(0.8, "cm"),
      legend.spacing.x = unit(0.4, "cm")
    )
  )
# ── Save ──────────────────────────────────────────────────────────────────────
dir.create("../figures/supplementary/spearman", showWarnings = FALSE, recursive = TRUE)

png("../figures/supplementary/spearman/Shannon_vs_NMFs_log.png",
    width = 10, height = 10, units = "in", res = 300)
print(combined_plot)
dev.off()

png("../figures/supplementary/spearman/Shannon_vs_NMFs_no_protein_log.png",
    width = 10, height = 10, units = "in", res = 300)
print(combined_plot_wo)
dev.off()

svg("../figures/supplementary/spearman/Shannon_vs_NMFs_log.svg",
    width = 10, height = 10)
print(combined_plot)
dev.off()

svg("../figures/supplementary/spearman/Shannon_vs_NMFs_no_protein_log.svg",
    width = 10, height = 10)
print(combined_plot_wo)
dev.off()
```

#Figure 4

```{r}
library(dplyr)
library(lme4)
library(broom.mixed)
library(performance)
library(ggplot2)

# ── Amino acid definitions ────────────────────────────────────────────────────
amino_acids <- c("Asp", "Glu", "hyP", "Ser", "Asn", "Gly", "Gln", "Thr", "Ala", "Tau",
                 "His", "Cit", "Pro", "Arg", "Val", "Tyr", "Met", "Cys2", "Ile", "Leu",
                 "Orn", "Lys", "Phe", "Trp", "Total_AAs", "Urea", "Creatinine", "UCA",
                 "LA", "PCA", "D.glucosamine", "Total_org._NMF")
amino_acids_wo_protein     <- paste0(amino_acids, "_wo_Protein")
Shiseido_metadata$Total_AAs_wo_Protein <- Shiseido_metadata$Total_Aas_wo_Protein

y_axis_labels_wo_protein <- c(
  "Asp_wo_Protein"            = "Aspartic Acid",
  "Glu_wo_Protein"            = "Glutamic Acid",
  "hyP_wo_Protein"            = "Hydroxyproline",
  "Ser_wo_Protein"            = "Serine",
  "Asn_wo_Protein"            = "Asparagine",
  "Gly_wo_Protein"            = "Glycine",
  "Gln_wo_Protein"            = "Glutamine",
  "Thr_wo_Protein"            = "Threonine",
  "Ala_wo_Protein"            = "Alanine",
  "Tau_wo_Protein"            = "Taurine",
  "His_wo_Protein"            = "Histidine",
  "Cit_wo_Protein"            = "Citrulline",
  "Pro_wo_Protein"            = "Proline",
  "Arg_wo_Protein"            = "Arginine",
  "Val_wo_Protein"            = "Valine",
  "Tyr_wo_Protein"            = "Tyrosine",
  "Met_wo_Protein"            = "Methionine",
  "Cys2_wo_Protein"           = "Cystine",
  "Ile_wo_Protein"            = "Isoleucine",
  "Leu_wo_Protein"            = "Leucine",
  "Orn_wo_Protein"            = "Ornithine",
  "Lys_wo_Protein"            = "Lysine",
  "Phe_wo_Protein"            = "Phenylalanine",
  "Trp_wo_Protein"            = "Tryptophan",
  "Total_AAs_wo_Protein"      = "Total Amino Acids",
  "Urea_wo_Protein"           = "Urea",
  "Creatinine_wo_Protein"     = "Creatinine",
  "UCA_wo_Protein"            = "Urocanic Acid",
  "LA_wo_Protein"             = "Lactic Acid",
  "PCA_wo_Protein"            = "Pyrrolidone Carboxylic Acid",
  "D.glucosamine_wo_Protein"  = "D-Glucosamine",
  "Total_org._NMF_wo_Protein" = "Total Organic NMF"
)

# ── Shared helper: subset and prepare data ────────────────────────────────────
prep_sub_df <- function(x, y, metadata) {
  metadata %>%
    filter(Face_site %in% c("Forehead", "Nose", "Cheek"),
           !is.na(Subject_ID), Subject_ID != "na") %>%
    select(Subject_ID, Face_site, all_of(c(x, y))) %>%
    filter(!is.na(.data[[x]]), !is.na(.data[[y]])) %>%
    mutate(Subject_ID = as.factor(Subject_ID),
           Face_site  = as.factor(Face_site))
}

# ── Shared helper: finalize LMM result dataframe ──────────────────────────────
finalize_lmm_df <- function(results_list) {
  df <- do.call(rbind, results_list)
  df$q_value <- p.adjust(df$p_value, method = "fdr")
  df <- df %>%
    mutate(
      signed_logq = -sign(Beta) * log10(q_value),
      stars = case_when(
        q_value < 0.001 ~ "***",
        q_value < 0.01  ~ "**",
        q_value < 0.05  ~ "*",
        TRUE            ~ ""
      ),
      X_label = x_axis_labels[X],
      Y_label = y_axis_labels_wo_protein[Y]
    )
  df$X_label <- factor(df$X_label, levels = x_axis_labels[x_vars])
  df$Y_label <- factor(df$Y_label,
                       levels = rev(y_axis_labels_wo_protein[amino_acids_wo_protein]))
  df
}

# ── Loop 1: raw LMM ───────────────────────────────────────────────────────────
lmm_results_raw <- list()

for (x in x_vars) {
  for (y in amino_acids_wo_protein) {
    sub_df <- prep_sub_df(x, y, Shiseido_metadata)
    if (nrow(sub_df) < 5) next

    model <- tryCatch(
      lmer(as.formula(paste(y, "~", x, "+ Face_site + (1|Subject_ID)")), data = sub_df),
      error = function(e) NULL
    )
    if (is.null(model)) next

    effect_row <- broom.mixed::tidy(model) %>% filter(term == x)
    if (nrow(effect_row) != 1) next

    lmm_results_raw[[paste(x, y, sep = "_")]] <- data.frame(
      X = x, Y = y,
      Beta    = effect_row$estimate,
      p_value = effect_row$p.value
    )
  }
}

lmm_df_raw <- finalize_lmm_df(lmm_results_raw)

# ── Loop 2: AIC comparison (raw vs log) ───────────────────────────────────────
model_comparison_results <- list()

for (x in x_vars) {
  for (y in amino_acids_wo_protein) {
    sub_df <- prep_sub_df(x, y, Shiseido_metadata)
    if (nrow(sub_df) < 5) next

    model_raw <- tryCatch(
      lmer(as.formula(paste(y, "~", x, "+ Face_site + (1|Subject_ID)")), data = sub_df),
      error = function(e) NULL
    )
    sub_df$log_x <- log10(pmax(as.numeric(sub_df[[x]]), 1))
    sub_df$log_y <- log10(pmax(as.numeric(sub_df[[y]]), 1))
    model_log <- tryCatch(
      lmer(log_y ~ log_x + Face_site + (1|Subject_ID), data = sub_df),
      error = function(e) NULL
    )

    if (is.null(model_raw) || is.null(model_log)) next

    delta_aic <- AIC(model_raw) - AIC(model_log)
    model_comparison_results[[paste(x, y, sep = "_")]] <- data.frame(
      X = x, Y = y,
      AIC_raw      = AIC(model_raw),
      AIC_log      = AIC(model_log),
      Delta_AIC    = delta_aic,
      Better_Model = ifelse(delta_aic > 2, "Log-transformed", "Untransformed or Similar")
    )
  }
}

model_comparison_df <- do.call(rbind, model_comparison_results)

# ── Loop 3: refit using best model per AIC ────────────────────────────────────
lmm_heatmap_results <- list()

for (i in seq_len(nrow(model_comparison_df))) {
  x            <- model_comparison_df$X[i]
  y            <- model_comparison_df$Y[i]
  better_model <- model_comparison_df$Better_Model[i]

  sub_df <- prep_sub_df(x, y, Shiseido_metadata)
  if (nrow(sub_df) < 5) next

  if (better_model == "Log-transformed") {
    sub_df$log_x <- log10(pmax(as.numeric(sub_df[[x]]), 1))
    sub_df$log_y <- log10(pmax(as.numeric(sub_df[[y]]), 1))
    model <- tryCatch(
      lmer(log_y ~ log_x + Face_site + (1|Subject_ID), data = sub_df),
      error = function(e) NULL
    )
    term <- "log_x"
  } else {
    model <- tryCatch(
      lmer(as.formula(paste(y, "~", x, "+ Face_site + (1|Subject_ID)")), data = sub_df),
      error = function(e) NULL
    )
    term <- x
  }

  if (is.null(model)) next
  effect_row <- broom.mixed::tidy(model) %>% filter(term == !!term)
  if (nrow(effect_row) != 1) next

  lmm_heatmap_results[[paste(x, y, sep = "_")]] <- data.frame(
    X = x, Y = y,
    Beta    = effect_row$estimate,
    p_value = effect_row$p.value
  )
}

lmm_df_heatmap <- finalize_lmm_df(lmm_heatmap_results)

# ── Shared heatmap theme ──────────────────────────────────────────────────────
heatmap_theme <- theme_minimal() +
  theme(
    axis.text.x   = element_text(angle = 45, hjust = 1),
    axis.text     = element_text(size = 12),
    axis.title    = element_text(size = 12),
    axis.ticks    = element_blank(),
    panel.border  = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.background = element_blank(),
    panel.grid    = element_blank()
  )

heatmap_fill <- scale_fill_gradient2(
  low      = "#4575b4",
  mid      = "white",
  high     = "#e08214",
  midpoint = 0,
  name     = expression(-sign(beta) * log[10](q))
)

# ── Heatmap helper ────────────────────────────────────────────────────────────
make_heatmap <- function(df, x_var, y_var, x_lab, y_lab, title) {
  ggplot(df, aes(x = .data[[x_var]], y = .data[[y_var]], fill = signed_logq)) +
    geom_tile(color = "white", width = 0.9, height = 0.9) +
    geom_text(aes(label = stars), color = "black", size = 5,
              hjust = 0.5, vjust = 0.8) +
    heatmap_fill +
    heatmap_theme +
    labs(x = x_lab, y = y_lab, title = title)
}

# ── Plot heatmaps ─────────────────────────────────────────────────────────────
heatmap_final <- make_heatmap(
  lmm_df_heatmap,
  x_var = "X_label", y_var = "Y_label",
  x_lab = NULL,
  y_lab = "Amino Acids and NMFs (pmol/cm²)",
  title = "LMM Associations (Best Model Fit)"
)

heatmap_transposed <- make_heatmap(
  lmm_df_heatmap,
  x_var = "Y_label", y_var = "X_label",
  x_lab = "Amino Acids and NMFs (pmol/cm²)",
  y_lab = NULL,
  title = "Associations between skin biophysical properties and targeted metabolites"
)

print(heatmap_final)
print(heatmap_transposed)
```



#### METABOLOMICS 

```{r packages}
library(tidyverse)
library(mixOmics)
library(ggpubr)
library(vegan)
library(caret)
library(rstatix)
library(UpSetR)
library(homologueDiscoverer)
library(ggdensity)
library(cowplot)
library(readxl)
library(lmerTest)
library(RColorBrewer)
library(MassTools)
library(shiny)
library(ggplot2)
library(dplyr)
library(stringr)
library(svglite)
library(Cairo)
```
```{r}
#Function for homologues analysis
plotAnnotatedStatic <- function(annotated, legend_setting = "bottom"){
  
  annotated <- mutate(annotated,
                      homologue_id = if_else(is.na(homologue_id), 0L, homologue_id),
                      homologue_id = as.factor(homologue_id))
  colourCount = length(unique(annotated$homologue_id))
  getPalette = colorRampPalette(RColorBrewer::brewer.pal(9, "Set1"))
  ncolor = length(unique(annotated$homologue_id))
  annotated <- arrange(annotated, desc(homologue_id))
  g <- ggplot(annotated, aes(group = homologue_id)) +
    geom_point(aes(x = rt, y = mz, color = homologue_id, size = homologue_id,
                   shape = homologue_id, alpha = homologue_id)) +
    geom_line(data = filter(annotated, homologue_id != 0),
              aes(x = rt, y = mz, group = homologue_id), alpha = 0.4, linewidth = 0.1) +
    ggtitle("Annotated Peak Table") +
    scale_colour_manual(values = c("black", getPalette(colourCount-1))) +
    scale_alpha_manual(values =  c(0.2, rep(0.85, times = ncolor))) +
    scale_size_manual(values = c(0.1, rep(1.5, times = ncolor))) +
    scale_shape_manual(values = c(1, rep(19, times = ncolor))) +
    xlab("Retention Time (s)") +
    ylab("Mass to Charge Ratio") +
    theme(legend.position=legend_setting, text = element_text(family="mono"))
  return(g)
  
}
```


```{r}
#Load in Data 
feature_table <- read_csv("../data/metabolomics/Britta_Shiseido_FBMN_2_attempt-d80187a83bd44a49901c57d813265012-featuretable_reformated.csv")


metadata <- read.csv("../data/Shiseido_metadata.txt", header=TRUE, sep="\t", stringsAsFactors=FALSE)
metadata <- metadata[metadata$X.SampleID != "" & !is.na(metadata$X.SampleID), ]


#sample_order <- read.csv("../Data/metabolomics/Run3_10252024/Sequence_order_10262024.csv")

#with suspects
#annotations <- read.delim("")
#without suspects
#annotations <- read.delim("../data/metabolomics/d80187a83bd44a49901c57d813265012-merged_results_with_gnps.tsv")
#annotations from run Britta_Shiseido_FBMN_2_attempt3quant_SZ_MC_PD_05072025
annotations <- read.delim("../data/metabolomics/Britta_Shiseido_FBMN_2_attempt3quant_SZ_MC_PD_05072025-merged_results_with_gnps.tsv")
annotations$X.Scan. <- as.character(annotations$X.Scan.)


canopus <- read_tsv("../data/metabolomics/canopus_structure_summary.tsv")
canopus$featureId <- as.character(canopus$mappingFeatureId)

mapping_file <- read.delim("../data/metabolomics/gnps_metadata.tsv")

```


```{r}
duplicates_in_metadata <- metadata[metadata$X.SampleID %in% metadata$X.SampleID[duplicated(metadata$X.SampleID)], ]

```

```{r}

metadata_2 <- metadata %>% full_join(mapping_file, by = c("ID" = "Study_SubjectIdentifierAsRecorded"))

```
```{r}
metadata_with_runorder <- metadata_2 %>%
  mutate(Order = as.numeric(str_extract(filename, "(?<=_)[0-9]+(?=\\.mzXML)"))) %>%
  arrange(Order)

```



```{r}
# metadata_with_runorder$ATTRIBUTE_Curated_SampleType_Sub1 contains the sample type column
# Show rows where 'ATTRIBUTE_Curated_SampleType_Sub1' equals the target value
filtered_rows <- metadata_with_runorder %>%
  filter(ATTRIBUTE_Curated_SampleType_Sub1 == "blank_analysis")

# Print the full rows
print(filtered_rows)

```

```{r}
### Import exported dataframe
Shiseidodf <- read.csv("../data/Shiseidodfc.txt", header=TRUE, sep="\t", stringsAsFactors=FALSE)
```
```{r}
### Order Face_site
Shiseidodf$Face_site<-factor(Shiseidodf$Face_site, levels=c("Forehead", "Nose", "Cheek", "Temple","Chin","Ear","Neck"))
```
```{r}
metadata_with_runorder2 <- metadata_with_runorder
```

```{r}
# Merge only Face_site for matching X.SampleID in Shiseido_metadata
metadata_with_runorder <- merge(
  metadata_with_runorder,
  Shiseidodf[, c("X.SampleID", "Face_site")],
  by = "X.SampleID",
  all.x = TRUE  # ensures only SampleIDs present in Shiseido_metadata are kept
)
```

```{r}
sum(metadata_with_runorder$ATTRIBUTE_Curated_SampleType_Sub1 == "skin", na.rm = TRUE)
```


```{r}
sum(metadata_with_runorder$ATTRIBUTE_Curated_SampleType_Sub1 == "blank_analysis", na.rm = TRUE)
```
```{r}
sum(metadata_with_runorder$ATTRIBUTE_Curated_SampleType_Sub1 == "reference material_personalcareproduct", na.rm = TRUE)
```

```{r}
info_feature <- feature_table %>% dplyr::select(1:3,8)
colnames(info_feature) <- c("Feature", "mz", "RT", "Corr_ID")
info_feature$Feature <- as.character(info_feature$Feature)

canopus$mappingFeatureId <-as.character(canopus$mappingFeatureId)
info_feature_complete <- info_feature %>% 
  left_join(annotations, by = c("Feature" = "X.Scan.")) %>%
  dplyr::select(1:4,18,24) %>% 
  left_join(canopus %>% distinct(mappingFeatureId, .keep_all = TRUE), by = c("Feature" = "mappingFeatureId"))
```




```{r}
metadata_with_runorder$filename_2 <- metadata_with_runorder$filename

metadata_with_runorder$filename_2 <- gsub(".mzXML", "", metadata_with_runorder$filename_2)
```


```{r}
# Extract data table 
data <- feature_table %>% 
  column_to_rownames("row ID") %>% dplyr::select(contains("Peak")) %>% 
  t() %>% as.data.frame() %>% rownames_to_column("SampleID") %>% 
  arrange(SampleID) %>% distinct(SampleID, .keep_all = TRUE)

data$SampleID <- gsub(".mzXML Peak area", "", data$SampleID)


```

```{r}
metadata_with_runorder$Subject_ID <- sub("-.*", "", metadata_with_runorder$microbiome_ID)

metadata_with_runorder_clean <- metadata_with_runorder %>%
  filter(!is.na(Subject_ID), !is.na(sebum))
```


```{r}
# List of SampleIDs in `data` that are not found in metadata
missing_samples <- data$SampleID[!data$SampleID %in% metadata_with_runorder$filename_2]

# Print missing sample IDs
print(missing_samples)

```

```{r}
# Samples in metadata but not in data
missing_in_data <- metadata_with_runorder$filename_2[!metadata_with_runorder$filename_2 %in% data$SampleID]

# Print them
print(missing_in_data)

```

```{r}
metadata_with_runorder <- metadata_with_runorder %>%
  filter(!is.na(filename_2))

```
```{r}
# Create and bind new rows with filename_2 set, all other columns as NA
metadata_with_runorder <- bind_rows(
  metadata_with_runorder,
  tibble(filename_2 = missing_samples)
)

```

```{r}
metadata_with_runorder <- metadata_with_runorder %>%
  mutate(Order = as.numeric(str_extract(filename_2, "\\d+(?=\\s*$|\\.mzXML$)")))

```

```{r}
metadata_with_runorder <- metadata_with_runorder %>%
  mutate(Sample_Type2 = case_when(
    ATTRIBUTE_Curated_SampleType_Sub1 == "skin" ~ "samples",
    ATTRIBUTE_Curated_SampleType_Sub1 == "reference material_personalcareproduct" ~ "BP",
    ATTRIBUTE_Curated_SampleType_Sub1 == "blank_analysis" ~ "blank_swabs",
    TRUE ~ NA_character_  # set the rest to NA for now
  )) %>%
  mutate(Sample_Type2 = case_when(
    !is.na(Sample_Type2) ~ Sample_Type2,  # keep already assigned
    str_detect(filename_2, regex("QC", ignore_case = TRUE)) ~ "QC",
    str_detect(filename_2, regex("wash", ignore_case = TRUE)) ~ "wash",
    str_detect(filename_2, regex("blank", ignore_case = TRUE)) ~ "blank",
    TRUE ~ NA_character_
  ))
```

```{r}
data_final <- data
```


```{r}
# Calculate Total Ion Count (TIC)


data_TIC <- data.frame(TIC = rowSums(data_final %>% column_to_rownames("SampleID"))) %>% 
  rownames_to_column("SampleID") %>%
  left_join(
    metadata_with_runorder %>% select(filename_2, Order, Subject_ID),
    by = c("SampleID" = "filename_2")
  )


```




```{r}

data_TIC2 <- data_TIC %>%
  filter(!str_detect(SampleID, regex("blank|wash|QC|emu|loit", ignore_case = TRUE)))

data_TIC %>%
  filter(!str_detect(SampleID, regex("blank|wash|QC|emu|loit", ignore_case = TRUE))) %>%
  #left_join(metadata_with_runorder, by = c("SampleID" = "filename_2")) %>%
  mutate(SubjectNo = as.factor(Subject_ID)) %>%
  ggscatter(x = "Order", y = "TIC", color = "Subject_ID", add = "reg.line") +
  stat_cor()


```



```{r}
# Check sample type
sample_tic <-  data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("blank|wash|QC|emu|loit", ignore_case = TRUE)))) %>% summarise(mean(TIC))
qc_tic <- data_TIC %>%   dplyr::filter(str_detect(SampleID, "QC") | str_detect(SampleID, regex("QC", ignore_case = TRUE))) %>% summarise(mean(TIC))

#blank swabs
blank_swabs_ids <- metadata_with_runorder %>%
  filter(Sample_Type2 == "blank_swabs") %>%
  pull(filename_2)

blank_swabs_tic <- data_TIC %>%
  filter(SampleID %in% blank_swabs_ids) %>%
  summarise(mean_TIC = mean(TIC, na.rm = TRUE))

#blank samples 

blank_ids <- metadata_with_runorder %>%
  filter(Sample_Type2 == "blank") %>%
  pull(filename_2)

blank_tic <- data_TIC %>%
  filter(SampleID %in% blank_ids) %>%
  summarise(mean_TIC = mean(TIC, na.rm = TRUE))

#BP 

BP_ids <- metadata_with_runorder %>%
  filter(Sample_Type2 == "BP") %>%
  pull(filename_2)

BP_tic <- data_TIC %>%
  filter(SampleID %in% BP_ids) %>%
  summarise(mean_TIC = mean(TIC, na.rm = TRUE))

#wash samples 

wash_ids <- metadata_with_runorder %>%
  filter(Sample_Type2 == "wash") %>%
  pull(filename_2)

wash_tic <- data_TIC %>%
  filter(SampleID %in% wash_ids) %>%
  summarise(mean_TIC = mean(TIC, na.rm = TRUE))



ratio_tic_sb <- sample_tic/blank_tic
ratio_tic_6p <- sample_tic/qc_tic
ratio_tic_swab <- sample_tic/blank_swabs_tic
ratio_tic_bp <- sample_tic/BP_tic

```

```{r}
a0 <- data_TIC %>% dplyr::filter(!(str_detect(SampleID, regex("blank|wash|QC|emu|loit", ignore_case = TRUE)))) %>% 
  ggscatter("Order", "TIC", add = "reg.line") +
  stat_cor() +ylim(0, 1e7) + ggtitle("TIC over Time for Samples")
```




```{r}
# Check internal standard and remove sample where there was a shift
fbmn_IS <- annotations %>% dplyr::filter(str_detect(Compound_Name, regex("sulf", ignore_case = TRUE))) %>% 
  distinct(X.Scan., .keep_all = TRUE) %>% dplyr::filter(Organism != "BILELIB19")
```
#Sulfadimethoxine
```{r}
#sulfadimethoxine was the one that was used according to Jasmine 
#X.Scan. 642
table_IS <- data_final %>% column_to_rownames("SampleID") %>% t() %>% as.data.frame() %>% rownames_to_column("ID") %>% 
  dplyr::filter(ID %in% fbmn_IS$X.Scan.) %>% column_to_rownames("ID") %>% t() %>% as.data.frame() %>% 
  rownames_to_column("filename") %>% left_join(metadata_with_runorder, by = c("filename" = "filename_2")) %>%
  dplyr::filter(!(str_detect(filename, regex("blank|wash|QC|emu|loit", ignore_case = TRUE)))) %>%
  dplyr::select(filename, `642`, Order, filename)
```

```{r}
colnames(table_IS)[2] <- "Sulfadimethoxine"

sulfadimethoxine <-table_IS %>% ggscatter(x = "Order", y = "Sulfadimethoxine", add = c("reg.line")) +
  stat_cor() + ylim(0, 5e4)
```
```{r}
ggsave("./Figures/Metabolomics/QC/Sulfadimethoxine.png", plot = last_plot(), dpi = 300, width = 8, height = 6, units = "in")
```

```{r}
table_IS %>% ggbarplot(x = "Order", y = "Sulfadimethoxine", xlab = "Run Order", 
                       ylab = "Peak Area sulfadimethoxine", title = "Internal Standard Acquisition") +
  geom_hline(yintercept = mean(table_IS$Sulfadimethoxine, na.rm = TRUE), linetype = "dashed", color = "blue")

cv_is <- sd(table_IS$Sulfadimethoxine)/mean(table_IS$Sulfadimethoxine)

#CV is 18,18452% 
#CV<30% is considered as good
```
```{r}
#sulfadimethoxine was the one that was used according to Jasmine 
#X.Scan. 222
table_IS <- data_final %>% column_to_rownames("SampleID") %>% t() %>% as.data.frame() %>% rownames_to_column("ID") %>% 
  dplyr::filter(ID %in% fbmn_IS$X.Scan.) %>% column_to_rownames("ID") %>% t() %>% as.data.frame() %>% 
  rownames_to_column("filename") %>% left_join(metadata_with_runorder, by = c("filename" = "filename_2")) %>%
  dplyr::filter(!(str_detect(filename, regex("blank|wash|QC|emu|loit", ignore_case = TRUE)))) %>%
  dplyr::select(filename, `222`, Order, filename)
```

```{r}
colnames(table_IS)[2] <- "Sulfachloropyridazine"

table_IS %>% ggscatter(x = "Order", y = "Sulfachloropyridazine", add = c("reg.line")) +
  stat_cor() + ylim(0, 5e3)
```
```{r}
table_IS %>% ggbarplot(x = "Order", y = "Sulfachloropyridazine", xlab = "Run Order", 
                       ylab = "Peak Area sulfachloropyridazine", title = "Internal Standard Acquisition") +
  geom_hline(yintercept = mean(table_IS$Sulfadimethoxine, na.rm = TRUE), linetype = "dashed", color = "blue")

cv_is <- sd(table_IS$Sulfachloropyridazine)/mean(table_IS$Sulfachloropyridazine)

#CV is 31,4125144% 
#CV<30% is considered as good
```



```{r}
# Check features in blanks, blank swabs, qc, and BP with filtering out the outliers
data_blank <- data_final %>%
  semi_join(
    metadata_with_runorder %>% filter(Sample_Type2 == "blank"),
    by = c("SampleID" = "filename_2")
  )

data_qc <- data_final %>%
  semi_join(
    metadata_with_runorder %>% filter(Sample_Type2 == "QC"),
    by = c("SampleID" = "filename_2")
  )

data_swabs <- data_final %>%
  semi_join(
    metadata_with_runorder %>% filter(Sample_Type2 == "blank_swabs"),
    by = c("SampleID" = "filename_2")
  )

data_BP <- data_final %>%
  semi_join(
    metadata_with_runorder %>% filter(Sample_Type2 == "BP"),
    by = c("SampleID" = "filename_2")
  )

data_skin <- data_final %>%
  semi_join(
    metadata_with_runorder %>% filter(Sample_Type2 == "samples"),
    by = c("SampleID" = "filename_2")
  )
```


```{r}
# Blanks
blanks_feature_info <- data.frame(Feature = colnames(data_blank)[-1],
                                  Mean_blank = data_blank %>% column_to_rownames("SampleID") %>% colMeans(), 
                                  SD_blank =  data_blank %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_blank = SD_blank/Mean_blank) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_blank, SD_blank, CV_blank) %>% 
  dplyr::filter(Mean_blank > 0) %>% arrange(desc(Mean_blank))

```




```{r}
# Six mix
qc_feature_info <- data.frame(Feature = colnames(data_qc)[-1],
                                  Mean_qc = data_qc %>% column_to_rownames("SampleID") %>% colMeans(), 
                                  SD_qc = data_qc %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_qc = SD_qc/Mean_qc) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_qc, SD_qc, CV_qc) %>% 
  dplyr::filter(Mean_qc > 0) %>% arrange(desc(Mean_qc))
```
```{r}
# CTLs
blankswabs_feature_info <- data.frame(Feature = colnames(data_swabs)[-1],
                                 Mean_ctl = data_swabs %>% column_to_rownames("SampleID") %>% colMeans(), 
                                 SD_ctl =  data_swabs %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_ctl = SD_ctl/Mean_ctl) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_ctl, SD_ctl, CV_ctl) %>% 
  dplyr::filter(Mean_ctl > 0) %>% arrange(desc(Mean_ctl))
```



```{r}
# BP
BP_feature_info <- data.frame(Feature = colnames(data_BP)[-1],
                                 Mean_BP = data_BP %>% column_to_rownames("SampleID") %>% colMeans(), 
                                 SD_BP =  data_BP %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_BP = SD_BP/Mean_BP) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_BP, SD_BP, CV_BP) %>% 
  dplyr::filter(Mean_BP > 0) %>% arrange(desc(Mean_BP))

```


```{r}
# Skin
skin_feature_info <- data.frame(Feature = colnames(data_skin)[-1],
                                 Mean_skin = data_skin %>% column_to_rownames("SampleID") %>% colMeans(), 
                                 SD_skin =  data_skin %>% column_to_rownames("SampleID") %>% apply(2, sd)) %>%
  dplyr::mutate(CV_skin = SD_skin/Mean_skin) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID, 
                                            Precursor_MZ, Mean_skin, SD_skin, CV_skin) %>% 
  dplyr::filter(Mean_skin > 0) %>% arrange(desc(Mean_skin))
```



```{r}
# Features to be removed Skin/Blank < 5 or 10
feature_to_remove <- blanks_feature_info %>% left_join(skin_feature_info) %>%
  dplyr::filter(Mean_blank > 0) %>% 
  dplyr::mutate(skin_Blank = Mean_skin/Mean_blank) %>% 
  dplyr::filter(skin_Blank < 5 | is.na(skin_Blank))

#42 features that need to be removed
```
```{r}
# Data with blank removal
data_clean <- data_final %>% dplyr::select(-c(feature_to_remove$Feature)) %>%
  column_to_rownames("SampleID") %>%
  select_if(~sum(.) != 0) %>% rownames_to_column("SampleID")
```

```{r}
# Features to be removed Skin/QCmix < 5 or 10 
feature_to_remove_mix <- qc_feature_info %>% left_join(skin_feature_info) %>% 
  dplyr::filter(Mean_qc > 0) %>% 
  dplyr::mutate(Skin_Mix = Mean_skin/Mean_qc) %>% 
  dplyr::filter(Skin_Mix < 5 | is.na(Skin_Mix)) %>% dplyr::filter(!(Feature %in% feature_to_remove$Feature))

#28 features 
```

```{r}
# Data with QC mix removal
data_clean2 <- data_clean %>% dplyr::select(-c(feature_to_remove_mix$Feature))

```
```{r}
# Features to be removed Skin/Skin_swabs < 5 or 10 
feature_to_remove_blankswabs <- blankswabs_feature_info %>% left_join(skin_feature_info) %>% 
  dplyr::filter(Mean_ctl > 0) %>% 
  dplyr::mutate(Skin_CTL = Mean_skin/Mean_ctl) %>% 
  dplyr::filter(Skin_CTL < 5 | is.na(Skin_CTL)) %>% dplyr::filter(!(Feature %in% feature_to_remove_mix$Feature))  %>%  dplyr::filter(!(Feature %in% feature_to_remove$Feature)) #%>% dplyr::filter(!(Feature %in% feature_to_remove_BP$Feature))
```


```{r}
# Data with removal of blank swabs
data_clean3 <- data_clean2 %>% dplyr::select(-c(feature_to_remove_blankswabs$Feature))


```

```{r}
# Features to be removed Skin/BP < 5 or 10 
feature_to_remove_BP <- BP_feature_info %>% left_join(skin_feature_info) %>% 
  dplyr::filter(Mean_BP > 0) %>% 
  dplyr::mutate(Skin_BP = Mean_skin/Mean_BP) %>% 
  dplyr::filter(Skin_BP < 5 | is.na(Skin_BP)) %>% dplyr::filter(!(Feature %in% feature_to_remove_mix$Feature))  %>%  dplyr::filter(!(Feature %in% feature_to_remove$Feature)) %>% dplyr::filter(!(Feature %in% feature_to_remove_blankswabs$Feature))

#868
```

```{r}
# Data with BP  removal
data_clean4 <- data_clean3 %>% dplyr::select(-c(feature_to_remove_BP$Feature))

```


```{r}
library(dplyr)

# Add label for each group of features
feature_to_remove_labeled <- feature_to_remove %>%
  mutate(Contaminant_Source = "blank")

feature_to_remove_mix_labeled <- feature_to_remove_mix %>%
  mutate(Contaminant_Source = "QC_mix")

feature_to_remove_blankswabs_labeled <- feature_to_remove_blankswabs %>%
  mutate(Contaminant_Source = "blank_swabs")

feature_to_remove_BP_labeled <- feature_to_remove_BP %>%
  mutate(Contaminant_Source = "BP")

```

```{r}
# Combine all into one table
features_to_remove_all <- bind_rows(
  feature_to_remove_labeled,
  feature_to_remove_mix_labeled,
  feature_to_remove_blankswabs_labeled,
  feature_to_remove_BP_labeled
) %>%
  distinct(Feature, Contaminant_Source)  # In case a feature appeared in multiple (rare, but safe to deduplicate)

```
```{r}
# Find features that remained (after all cleaning)
remaining_features <- colnames(data_clean4)
remaining_features <- remaining_features[remaining_features != "SampleID"]  # Remove SampleID column if still present

# Create a dataframe for skin features
skin_features_df <- data.frame(
  Feature = remaining_features,
  Contaminant_Source = "skin"
)

```

```{r}
# Combine removed features (blanks, QC mix, blank swabs, BP) and remaining skin features
final_feature_source_table <- bind_rows(
  features_to_remove_all,
  skin_features_df
)

```

```{r}
# Define the color palette for six Sample_Type categories
sample_type_colors <- c("samples" = "#daa600",  # Blue
                        "blank" = "#2ca25f",  # Orange
                        "BP" = "#c00001",
                        "QC" = "#bcbddc",  # Red
                        "blank_swabs" = "#247cb9",  # Purple
                        "wash" = "#8c564b")  # Brown
```


```{r}
# PCA raw data before removing the features
PCA_raw <- mixOmics::pca(data_final %>% select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>%
                           column_to_rownames("SampleID"), ncomp = 2, center = TRUE, scale = TRUE)
PCA_raw_scores <- data.frame(PCA_raw$variates$X) %>% rownames_to_column("SampleID") %>% left_join(metadata_with_runorder, by = c("SampleID" = "filename_2"))

i <- "Sample_Type2"

PCA_raw_plot <- PCA_raw_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA of Raw Metabolomics Data by", i, sep = " "),
            xlab = paste("PC1 (", round(PCA_raw$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_raw$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_raw_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed() +
  scale_color_manual(values = sample_type_colors)

```

```{r}
# 3. Save the full layout
ggsave("../figures/metabolomics/qc/PCA_raw_plot_SampleType2.png",
       plot = PCA_raw_plot,
       width = 5, height = 5, dpi = 300)

ggsave("../figures/metabolomics/qc/PCA_raw_plot_SampleType2.svg",
       plot = PCA_raw_plot,
       width = 5, height = 5)
```

```{r}
data_final_filtered <- data_final %>%
  column_to_rownames("SampleID") %>%
  filter(rowSums(.) > 0)
```


```{r}
# Apply rclr
data_final_clr <- decostand(data_final_filtered, method = "rclr")
```

```{r}
PCA_raw <- mixOmics::pca(data_final_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_raw_scores <- data.frame(PCA_raw$variates$X) %>% rownames_to_column("SampleID") %>% left_join(metadata_with_runorder, by = c("SampleID" = "filename_2"))

i <- "Sample_Type2"

PCA_raw_plot <- PCA_raw_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA of rCLR-Transformed Metabolomics Data by", i, sep = " "),
            xlab = paste("PC1 (", round(PCA_raw$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_raw$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_raw_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed() +
  scale_color_manual(values = sample_type_colors)
```

```{r}
# 3. Save the full layout
ggsave("../figures/metabolomics/qc/PCA_raw_plot_SampleType2_rclr.png",
       plot = PCA_raw_plot,
       width = 5, height = 5, dpi = 300)

ggsave("../figures/metabolomics/qc/PCA_raw_plot_SampleType2_rclr.svg",
       plot = PCA_raw_plot,
       width = 5, height = 5)
```



```{r}
# Remove rows with all zeros BEFORE rclr
data_filtered <- data_clean4 %>%
  column_to_rownames("SampleID") %>%
  filter(rowSums(.) > 0)

# Apply rclr
data_clean_4_clr <- decostand(data_filtered, method = "rclr")
```


```{r}
# PCA
PCA_whole <- mixOmics::pca(data_clean_4_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID")  %>% left_join(metadata_with_runorder, by = c("SampleID" = "filename_2"))
#PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% rownames_to_column("SampleID") %>% left_join(metadata_final)
i <- "Sample_Type2"

PCA_plot <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Skin Metabolome - rclr transformed data"),
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()  +
  scale_color_manual(values = sample_type_colors)
```


```{r}
# Remove plastic contamination
feature_check <- info_feature %>% dplyr::select(1:3) %>% 
  dplyr::filter(Feature %in% colnames(data_clean4)) %>% as_tibble()
colnames(feature_check) <- c("peak_id", "mz", "rt")
feature_check$mz <- as.double(feature_check$mz)
feature_check$rt <- as.double(feature_check$rt)
```

```{r}
# for 1 CTL sample 
feature_pool <- data_clean4 %>% 
  dplyr::filter(SampleID == "blank_GG11_01_1689") %>% column_to_rownames("SampleID") %>%
  t() %>% as.data.frame() %>% rownames_to_column("peak_id")

feature_check_pool <- feature_check %>% left_join(feature_pool)

feature_check_pool$peak_id <- as.integer(feature_check_pool$peak_id)
colnames(feature_check_pool)[4] <- "intensity"
```

```{r}
# Remove PEGs
peak_table_PEG <- detectHomologues(feature_check_pool, mz_steps = c(44.02628),
                                   rt_min = 0.1, rt_max = 100, ppm_tolerance = 10, 
                                   min_series_length = 5, search_mode = "targeted", 
                                   step_mode = "increment", verbose = TRUE)

plotAnnotatedStatic(peak_table_PEG)
```

```{r}

sdb <- sdbCreate(peak_table_PEG, sample_origin = "data")


```

```{r}
# Remove untargeted 
peak_table_untargeted <- detectHomologues(feature_check_pool, mz_min = 10, mz_max = 50, 
                                          rt_min = 0.5, rt_max = 100, ppm_tolerance = 5, 
                                          min_series_length = 5, search_mode = "untargeted", 
                                          step_mode = "increment", verbose = TRUE)
```

```{r}
#if running targeted and untargeted 
# Combine tables
overlapping_ids <- sdbCheckContained(sdb, peak_table_untargeted, ppm_tolerance = 0.5, rt_tolerance = 0.5)
sdb <- sdbPush(sdb, filter(peak_table_untargeted, !(homologue_id %in% overlapping_ids)))
```

```{r}
sdb$sample_peak_id <- as.character(sdb$sample_peak_id)
features_contaminat <- info_feature_complete %>% dplyr::filter(Feature %in% sdb$sample_peak_id)

data_clean5 <- data_clean4 %>% dplyr::select(-c(sdb$sample_peak_id))

```


```{r}
# PCA
PCA_raw <- mixOmics::pca(data_clean5 %>% select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>%
                           column_to_rownames("SampleID"), ncomp = 2, center = TRUE, scale = TRUE)
PCA_raw_scores <- data.frame(PCA_raw$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_with_runorder, by = c("SampleID" = "filename_2"))

i <- "Sample_Type2"

PCA_raw_plot <- PCA_raw_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = paste("PCA - Skin Metabolome", i, sep = " "),
            xlab = paste("PC1 (", round(PCA_raw$prop_expl_var$X[1]*100, digits = 1),"%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_raw$prop_expl_var$X[2]*100, digits = 1),"%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_raw_scores %>% group_by((!!sym(i))) %>% 
               summarise_at(vars(matches("PC")), mean), aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 10), axis.title = element_text(size = 8),
        axis.text = element_text(size = 4)) + coord_fixed()+
  scale_color_manual(values = sample_type_colors)

```

```{r}
# Keep only samples
data_sample <- data_clean5 %>% 
   dplyr::filter(!(str_detect(SampleID, regex("blank|wash|QC|emu|loit", ignore_case = TRUE)))) %>% 
  column_to_rownames("SampleID") %>%
  filter(rowSums(.) > 0)%>%  
  select_at(vars(-one_of(nearZeroVar(., names = TRUE)))) %>% rownames_to_column("SampleID") 
```

```{r}
# after filtering out the contaminant featuers
filtered_data_sample_skin_feature_info <- data.frame(Feature = colnames(data_sample)[-1]) %>% left_join(annotations, by = c("Feature" = "X.Scan.")) %>% 
  left_join(info_feature) %>% dplyr::select(Feature, mz, RT, Compound_Name, SpectrumID) 
```



```{r}
fill_colors <- c(
  "Forehead" = "#ffc65e",   # warm amber
  "Nose"     = "#5693c7",   # soft blue
  "Cheek"    = "#8ccd87",   # light green
  "Ear"      = "#9dcfcf",   # muted teal (cooler, pairs with Nose & Cheek)
  "Temple"   = "#f29d1e",   # bold warm orange (near Forehead)
  "Chin"     = "#bfa6d0",   # lavender (softer contrast after orange)
  "Neck"     = "#d3d3d3"    # neutral gray (least visually dominant, good for last)
)

border_colors <- c(
  "Forehead" = "#e0a846",  # darker amber
  "Nose"     = "#3a6b99",  # darker blue
  "Cheek"    = "#4a7a4f",  # deeper green
  "Ear"      = "#6ca5a5",  # darker teal
  "Temple"   = "#c97e0f",  # deeper warm orange
  "Chin"     = "#9989ac",  # deeper lavender
  "Neck"     = "#a9a9a9"   # medium gray
)
```


```{r}
data_sample_clr <- decostand(data_sample %>% column_to_rownames("SampleID"), method = "rclr")
```
```{r}
# PCA
PCA_whole <- mixOmics::pca(data_sample_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_with_runorder, by = c("SampleID" = "filename_2"))
PCA_whole_scores$Face_site <- factor(PCA_whole_scores$Face_site, levels=c("Forehead", "Nose", "Cheek", "Temple","Chin","Ear","Neck"))

i <- "Face_site"


```



```{r}
PCA_plot_Face_site <- PCA_whole_scores %>%
  ggplot(aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i))) +
  geom_point(shape = 21, size = 3, stroke = 0.6, alpha = 0.6) +
  geom_point(
    data = PCA_whole_scores %>%
      group_by((!!sym(i))) %>%
      summarise(across(matches("PC"), mean)),
    aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i)),
    shape = 8, size = 4, inherit.aes = FALSE
  ) +
  scale_fill_manual(name = "Face site", values = fill_colors) +  # set title here
  scale_color_manual(name = "Face site", values = border_colors) +  # match title
  labs(
    title = "PCA of RCLR-Transformed Metabolite Features by Face Site",
    x = paste0("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)"),
    y = paste0("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)")
  ) +
  coord_fixed() +
  theme_classic(base_size = 10) +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 15), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )

```


```{r}
# PERMANOVA
dist_metabolites <- vegdist(data_sample_clr, method = "euclidean")
disper_donor <- betadisper(dist_metabolites, PCA_whole_scores$Subject_ID)

#significant variation between the groups
anova(disper_donor)
permanova <- adonis2(dist_metabolites ~ PCA_whole_scores$Subject_ID + PCA_whole_scores$Face_site, PCA_whole_scores, na.action = na.omit, by="margin")
```

# plot with Permanova

```{r}
face_row <- permanova[2, ]
r2 <- round(face_row$R2 * 100, 1)
f_val <- round(face_row$F, 2)
p_val <- format(face_row$`Pr(>F)`, digits = 3, scientific = TRUE)

label_text <- paste0("PERMANOVA\nR² = ", r2, "%\nF = ", f_val, "\np = ", p_val)

```

```{r}
PCA_plot_Face_site <- PCA_whole_scores %>%
  ggplot(aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i))) +
  geom_point(shape = 21, size = 3, stroke = 0.6, alpha = 0.6) +
  geom_point(
    data = PCA_whole_scores %>%
      group_by((!!sym(i))) %>%
      summarise(across(matches("PC"), mean)),
    aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i)),
    shape = 8, size = 4, inherit.aes = FALSE
  ) +
  scale_fill_manual(name = "Face site", values = fill_colors) +  # set title here
  scale_color_manual(name = "Face site", values = border_colors) +  # match title
  labs(
    title = "PCA of RCLR-Transformed Untargeted Metabolite Features by Face Site", #PCA of RCLR-Transformed Metabolite Features by Face Site
    x = paste0("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)"),
    y = paste0("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)")
  ) +
  coord_fixed() +
  theme_classic(base_size = 10) +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 15), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )+ annotate("text", 
    x = Inf, y = Inf,
    label = label_text,
    hjust = 1.1, vjust = 1.2, size = 3.8
  )


```

# leave out Neck, Chin, Temple, Ear ONLY FACE NOSE CHEEK
```{r}
# First join metadata and filter to include only relevant face sites
metadata_filtered <- metadata_with_runorder %>%
  filter(Face_site %in% c("Forehead", "Nose", "Cheek"))

# Then subset the data_sample accordingly
data_sample_filtered <- data_sample %>%
  filter(SampleID %in% metadata_filtered$filename_2)

# CLR transformation on filtered data
data_sample_clr <- decostand(data_sample_filtered %>% column_to_rownames("SampleID"), method = "rclr")

# Run PCA
PCA_whole <- mixOmics::pca(data_sample_clr, ncomp = 2, center = TRUE, scale = TRUE)

# Join PCA scores with filtered metadata
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID") %>%
  left_join(metadata_filtered, by = c("SampleID" = "filename_2"))

# Set Face_site as ordered factor
PCA_whole_scores$Face_site <- factor(PCA_whole_scores$Face_site, levels = c("Forehead", "Nose", "Cheek"))

i <- "Face_site"


```


```{r}
PCA_plot_Face_site <- PCA_whole_scores %>%
  ggplot(aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i))) +
  geom_point(shape = 21, size = 3, stroke = 0.6, alpha = 0.6) +
  geom_point(
    data = PCA_whole_scores %>%
      group_by((!!sym(i))) %>%
      summarise(across(matches("PC"), mean)),
    aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i)),
    shape = 8, size = 4, inherit.aes = FALSE
  ) +
  scale_fill_manual(name = "Face site", values = fill_colors) +  # set title here
  scale_color_manual(name = "Face site", values = border_colors) +  # match title
  labs(
    title = "PCA of RCLR-Transformed Metabolite Features by Face Site",
    x = paste0("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)"),
    y = paste0("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)")
  ) +
  coord_fixed() +
  theme_classic(base_size = 10) +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 15), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  )

```

```{r}
# PERMANOVA
dist_metabolites <- vegdist(data_sample_clr, method = "euclidean")
disper_donor <- betadisper(dist_metabolites, PCA_whole_scores$Subject_ID)

#significant variation between the groups
anova(disper_donor)
permanova <- adonis2(dist_metabolites ~ PCA_whole_scores$Subject_ID + PCA_whole_scores$Face_site, PCA_whole_scores, na.action = na.omit, by="margin")
```

# plot with Permanova

```{r}
face_row <- permanova[2, ]
r2 <- round(face_row$R2 * 100, 1)
f_val <- round(face_row$F, 2)
p_val <- format(face_row$`Pr(>F)`, digits = 3, scientific = TRUE)

label_text <- paste0("PERMANOVA\nR² = ", r2, "%\nF = ", f_val, "\np = ", p_val)

```

```{r}
# 1. Create named vector for labels with counts
# 1. Create named vector for legend labels with Face_site as name and proper label
face_site_counts_df <- PCA_whole_scores %>%
  filter(!is.na(!!sym(i))) %>%
  count(face_site = !!sym(i)) %>%
  mutate(label = paste0(face_site, " (n = ", n, ")"))

# 2. Create a named vector: names = face_site, values = label
face_site_labels <- setNames(face_site_counts_df$label, face_site_counts_df$face_site)


PCA_whole_scores[[i]] <- factor(PCA_whole_scores[[i]], levels = names(face_site_labels))
# 3. Now update the plot with manual legend labels
PCA_plot_Face_site <- PCA_whole_scores %>%
  ggplot(aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i))) +
  geom_point(shape = 21, size = 3, stroke = 0.6, alpha = 0.6) +
  geom_point(
    data = PCA_whole_scores %>%
      group_by((!!sym(i))) %>%
      summarise(across(matches("PC"), mean), .groups = "drop"),
    aes(x = PC1, y = PC2, fill = !!sym(i), color = !!sym(i)),
    shape = 8, size = 4, inherit.aes = FALSE
  ) +
  scale_fill_manual(
    name = "Face site",
    values = fill_colors,
    labels = face_site_labels
  ) +
  scale_color_manual(
    name = "Face site",
    values = border_colors,
    labels = face_site_labels
  ) +
  labs(
    title = "PCA of RCLR-Transformed Untargeted Metabolite Features by Face Site",
    x = paste0("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)"),
    y = paste0("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)")
  ) +
  coord_fixed() +
  theme_classic(base_size = 10) +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 15), 
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  annotate("text", 
    x = Inf, y = Inf,
    label = label_text,
    hjust = 1.1, vjust = 1.2, size = 3.8
  )

```


```{r}

i <- "Subject_ID"  # change coloring to Subject_ID

PCA_plot_subject <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = "PCA of RCLR-Transformed Metabolite Features by Subject",
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, digits = 1), "%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, digits = 1), "%)", sep = ""),
            ggtheme = theme_classic()) +
  geom_point(data = PCA_whole_scores %>%
               group_by((!!sym(i))) %>%
               summarise(across(matches("PC"), mean)), 
             aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  theme(plot.title = element_text(size = 15), 
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 12)) +
  coord_fixed() + scale_color_viridis_d(option = "turbo")  # vibrant and high-contrast


```

```{r}
library(ggrepel)

i <- "Subject_ID"

PCA_plot_subject <- ggplot(PCA_whole_scores, aes(x = PC1, y = PC2, color = !!sym(i), group = !!sym(i))) +
  geom_point(alpha = 0.6) +
  geom_path(alpha = 0.5, linewidth = 0.6) +  # lines connecting points from same subject
  geom_text_repel(
    data = PCA_whole_scores %>%
      group_by((!!sym(i))) %>%
      summarise(across(c(PC1, PC2), mean), .groups = "drop"),
    aes(label = !!sym(i), x = PC1, y = PC2, color = !!sym(i)),
    size = 3, show.legend = FALSE, max.overlaps = Inf
  ) +
  scale_color_viridis_d(option = "turbo") +
  coord_fixed() +
  labs(
    title = "PCA of RCLR-Transformed Metabolites by Subject",
    x = paste0("PC1 (", round(PCA_whole$prop_expl_var$X[1] * 100, 1), "%)"),
    y = paste0("PC2 (", round(PCA_whole$prop_expl_var$X[2] * 100, 1), "%)")
  ) +
  theme_classic(base_size = 12) +
  theme(legend.position = "none")

```

```{r}
# 3. Save the full layout
ggsave("../figures/supplementary/suppl_fig10.png",
       plot = PCA_plot_subject,
       width = 9, height = 9, dpi = 300)

ggsave("../figures/supplementary/suppl_fig10.svg",
       plot = PCA_plot_subject,
       width = 9, height = 9)
```



```{r}
q <- quantile(metadata_with_runorder$sebum, probs = c(0.25, 0.75), na.rm = TRUE)
low_thresh <- q[1]
high_thresh <- q[2]

# 2. Assign sebum groups
metadata_with_runorder <- metadata_with_runorder %>%
  mutate(sebum_group = case_when(
    sebum < low_thresh ~ "low",
    sebum >= low_thresh & sebum <= high_thresh ~ "medium",
    sebum > high_thresh ~ "high"
  ))

```



```{r}
# RCLR transformation
data_sample_clr <- decostand(data_sample %>% column_to_rownames("SampleID"), method = "rclr")
```

```{r}
# PCA
PCA_whole <- mixOmics::pca(data_sample_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_with_runorder, by = c("SampleID" = "filename_2"))

```

```{r}
# Example thresholds based on quantiles
q <- quantile(metadata_with_runorder$sebum, probs = c(0.25, 0.75), na.rm = TRUE)
low_thresh <- round(q[1], 1)
high_thresh <- round(q[2], 1)

# Set factor levels manually so Low, Medium, High appear in the correct order
PCA_whole_scores$sebum_group <- factor(PCA_whole_scores$sebum_group, 
                                       levels = c("low", "medium", "high"))

# Set i
i <- "sebum_group"

# Create PCA plot
PCA_plot_sebum <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = "PCA of RCLR-Transformed Metabolite Features by Sebum Level",
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)", sep = ""),
            ggtheme = theme_classic()) +
  
  geom_point(data = PCA_whole_scores %>%
               group_by((!!sym(i))) %>%
               summarise(across(matches("PC"), mean)), 
             aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  
  scale_color_manual(
    name = "Sebum Group",
    values = c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214"),
    labels = c(
      low = paste0("Low (< ", low_thresh, " µg/cm²)"),
      medium = paste0("Medium (", low_thresh, "–", high_thresh, " µg/cm²)"),
      high = paste0("High (> ", high_thresh, " µg/cm²)")
    )
  ) +
  
  theme(plot.title = element_text(size = 15), 
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 11)) +
  
  coord_fixed()

```




```{r}
# PERMANOVA
dist_metabolites <- vegdist(data_sample_clr, method = "euclidean")
disper_donor <- betadisper(dist_metabolites, PCA_whole_scores$Subject_ID)

#significant variation between the groups
anova(disper_donor)
permanova_sebumgroup <- adonis2(dist_metabolites ~ PCA_whole_scores$Subject_ID + PCA_whole_scores$sebum_group, PCA_whole_scores, na.action = na.omit, by="margin")
```


```{r}
sebum_row <- permanova_sebumgroup[2, ]
r2 <- round(sebum_row$R2 * 100, 1)
f_val <- round(sebum_row$F, 2)
p_val <- format(sebum_row$`Pr(>F)`, digits = 3, scientific = TRUE)

label_text <- paste0("PERMANOVA\nR² = ", r2, "%\nF = ", f_val, "\np = ", p_val)

```

#Plot with Permanova

```{r}
# Example thresholds based on quantiles
q <- quantile(metadata_with_runorder$sebum, probs = c(0.25, 0.75), na.rm = TRUE)
low_thresh <- round(q[1], 1)
high_thresh <- round(q[2], 1)

# Set factor levels manually so Low, Medium, High appear in the correct order
PCA_whole_scores$sebum_group <- factor(PCA_whole_scores$sebum_group, 
                                       levels = c("low", "medium", "high"))

# Set i
i <- "sebum_group"

# Create PCA plot
PCA_plot_sebum <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = "PCA of RCLR-Transformed Untargeted Metabolite Features by Sebum Level", #"PCA of RCLR-Transformed Metabolite Features by Sebum Level"
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)", sep = ""),
            ggtheme = theme_classic()) +
  
  geom_point(data = PCA_whole_scores %>%
               group_by((!!sym(i))) %>%
               summarise(across(matches("PC"), mean)), 
             aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  
  scale_color_manual(
    name = "Sebum Group",
    values = c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214"),
    labels = c(
      low = paste0("Low (< ", low_thresh, " µg/cm²)"),
      medium = paste0("Medium (", low_thresh, "–", high_thresh, " µg/cm²)"),
      high = paste0("High (> ", high_thresh, " µg/cm²)")
    )
  ) +
  
  theme(plot.title = element_text(size = 15), 
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 11)) +
  
  coord_fixed()+ annotate("text", 
    x = Inf, y = Inf,
    label = label_text,
    hjust = 1.1, vjust = 1.2, size = 3.8
  )

```

#including counts per group

```{r}
# 2. Create legend labels with n + thresholds
sebum_group_counts_df <- PCA_whole_scores %>%
  filter(!is.na(sebum_group)) %>%
  count(sebum_group) %>%
  mutate(
    label = case_when(
      sebum_group == "low" ~ paste0("Low (< ", low_thresh, " µg/cm², n = ", n, ")"),
      sebum_group == "medium" ~ paste0("Medium (", low_thresh, "–", high_thresh, " µg/cm², n = ", n, ")"),
      sebum_group == "high" ~ paste0("High (> ", high_thresh, " µg/cm², n = ", n, ")"),
      TRUE ~ as.character(sebum_group)
    )
  )

# 3. Create named vector for legend
sebum_group_labels <- setNames(sebum_group_counts_df$label, sebum_group_counts_df$sebum_group)

# 4. Set factor level order
PCA_whole_scores$sebum_group <- factor(PCA_whole_scores$sebum_group, levels = c("low", "medium", "high"))

# 5. Generate PCA plot
PCA_plot_sebum <- PCA_whole_scores %>%
  ggplot(aes(x = PC1, y = PC2, fill = sebum_group, color = sebum_group)) +
  geom_point(shape = 21, size = 3, stroke = 0.6, alpha = 0.6) +
  geom_point(
    data = PCA_whole_scores %>%
      group_by(sebum_group) %>%
      summarise(across(matches("PC"), mean), .groups = "drop"),
    aes(x = PC1, y = PC2, fill = sebum_group, color = sebum_group),
    shape = 8, size = 4, inherit.aes = FALSE
  ) +
  scale_fill_manual(
    name = "Sebum Group",
    values = c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214"),
    labels = sebum_group_labels
  ) +
  scale_color_manual(
    name = "Sebum Group",
    values = c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214"),
    labels = sebum_group_labels
  ) +
  labs(
    title = "PCA of RCLR-Transformed Untargeted Metabolite Features by Sebum Level",
    x = paste0("PC1 (", round(PCA_whole$prop_expl_var$X[1] * 100, 1), "%)"),
    y = paste0("PC2 (", round(PCA_whole$prop_expl_var$X[2] * 100, 1), "%)")
  ) +
  coord_fixed() +
  theme_classic(base_size = 10) +
  theme(
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 11),
    plot.title = element_text(size = 15),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  annotate("text", 
    x = Inf, y = Inf,
    label = label_text,
    hjust = 1.1, vjust = 1.2, size = 3.8
  )

```

#Figure 4b
```{r}
features_skin <- skin_feature_info %>% filter(Mean_skin > 0) %>% pull(Feature)
features_BP <- BP_feature_info %>% filter(Mean_BP > 0) %>% pull(Feature)
features_QC <- qc_feature_info %>% filter(Mean_qc > 0) %>% pull(Feature)
features_blank <- blanks_feature_info %>% filter(Mean_blank > 0) %>% pull(Feature)
features_blankswabs <- blankswabs_feature_info %>% filter(Mean_ctl > 0) %>% pull(Feature)

```

```{r}


feature_lists <- list(
  Skin = features_skin,
  BP = features_BP,
  QC = features_QC,
  Blank = features_blank,
  "Blank Swabs" = features_blankswabs
)


```

```{r}
library(eulerr)

# Rename the lists

feature_lists_named <- feature_lists
names(feature_lists_named) <- c(
  "Skin",
  "Personal Care Ref",     # or whatever term you prefer
  "Quality Control",
  "Blank",
  "Blank Swabs"
)

# Fit Venn diagram
fit <- euler(feature_lists_named)

# Plot
venn_plot <- plot(fit,
     fills = c(
       "Skin" = "#daa600",
       "Personal Care Ref" = "#c00001",
       "Quality Control" = "#bcbddc",
       "Blank" = "#2ca25f",
       "Blank Swabs" = "#247cb9"
     ),
     #labels = list(font = 2),
     quantities = TRUE,
     legend = list(title = "Sample type", cex = 1),
     main = list(label = "Overlap of Untargeted Metabolite Features Across Sample Types", cex = 1.2)
)
#Overlap of Detected Features Across Sample Types
```

```{r}
venn_plot2 <- ggarrange(
  venn_plot, NULL,
  ncol = 2,
  nrow = 1,
  labels = c("b", "c"),
  common.legend = FALSE ,font.label = list(size = 18, face = "bold"), align ="hv"  # adjust size and style
)
```



```{r}
venn_pca <- ggarrange(
  PCA_plot_Face_site, PCA_plot_sebum, 
  ncol = 2,
  nrow = 1,
  labels = c("d", "e"),
  common.legend = FALSE ,font.label = list(size = 18, face = "bold"), align ="hv"  # adjust size and style
)
```

```{r}
top_grid <- ggarrange(heatmap_transposed,venn_plot2, venn_pca,
  ncol = 1, nrow = 3,
  labels = c("a", "", ""), # no label on boxplot_fg as it already has f/g inside
  heights = c(0.85,1, 1),font.label = list(size = 18, face = "bold")  # adjust size and style
  # optional: give a bit more room to the bottom row
)
```

```{r}
# 3. Save the full layout
ggsave("../figures/main/Figure4.png", #venn_sebum_3.png
       plot = top_grid,
       width = 16, height = 16, dpi = 300)

ggsave("../figures/main/Figure4.svg", #venn_sebum_3.svg
       plot = top_grid,
       width = 16, height = 16)
```



#Run PCA / PLSDA on only extreme cases of sebum concentration (7ug/cm^2 < and > 30 ug/cm^2); filtering of the table 
```{r}
# Filter and pull the filename2 values
extreme_filenames <- metadata_with_runorder_clean %>%
  dplyr::filter(sebum < 3.5 | sebum > 16.9) %>%  #7 and 20
  dplyr::pull(filename_2)
#Face_site == "Forehead", 
# View the result
extreme_filenames

```

```{r}
#ALTERNATIVE: FILTER OUT THOSE SAMPLES that are related to forehead, nose, cheek

extreme_filenames <- metadata_with_runorder_clean %>%
  dplyr::filter(
    (sebum < 3.5 | sebum > 16.9) &
    Face_site %in% c("Forehead", "Nose", "Cheek")
  ) %>%
  dplyr::pull(filename_2)

# View result
extreme_filenames

```

```{r}
# Filter data_sample to keep only the extreme samples
data_sample_extreme <- data_sample %>%
  dplyr::filter(SampleID %in% extreme_filenames)

# View the result
data_sample_extreme

```


```{r}
# First, create the new column based on sebum values
metadata_with_runorder_clean <- metadata_with_runorder_clean %>%
  dplyr::mutate(
    Extreme_Sebum = dplyr::case_when(
      filename_2 %in% data_sample_extreme$SampleID & sebum < 3.5 ~ "Low", #7
      filename_2 %in% data_sample_extreme$SampleID & sebum > 16.9 ~ "High", #20
      TRUE ~ NA_character_  # Keep NA for all others
    )
  )

# View result
metadata_with_runorder_clean

```
```{r}
# Count number of Low and High in Extreme_Sebum
extreme_sebum_counts <- metadata_with_runorder_clean %>%
  dplyr::filter(!is.na(Extreme_Sebum)) %>%
  dplyr::count(Extreme_Sebum)

# View result
extreme_sebum_counts

```
```{r}
# Assuming extreme_sebum_counts already exists
low_n <- extreme_sebum_counts %>% filter(Extreme_Sebum == "Low") %>% pull(n)
high_n <- extreme_sebum_counts %>% filter(Extreme_Sebum == "High") %>% pull(n)

```


```{r}
# RCLR transformation
data_sample_extreme_clr <- decostand(data_sample_extreme %>% column_to_rownames("SampleID"), method = "rclr")
```
```{r}
# PCA
PCA_whole <- mixOmics::pca(data_sample_extreme_clr, ncomp = 2, center = TRUE, scale = TRUE)
PCA_whole_scores <- data.frame(PCA_whole$variates$X) %>% 
  rownames_to_column("SampleID") %>% left_join(metadata_with_runorder_clean, by = c("SampleID" = "filename_2"))
#PCA_whole_scores$Face_site <- factor(PCA_whole_scores$Face_site, levels=c("Forehead", "Nose", "Cheek", "Temple","Chin","Ear","Neck"))

i <- "Extreme_Sebum"


```


#correct counts
```{r}
# Count number of Low and High in Extreme_Sebum
sebum_counts <- PCA_whole_scores %>%
  dplyr::filter(!is.na(Extreme_Sebum)) %>%
  dplyr::count(Extreme_Sebum)

# View result
sebum_counts
```

```{r}
# Assuming extreme_sebum_counts already exists
low_n <- sebum_counts %>% filter(Extreme_Sebum == "Low") %>% pull(n)
medium_n <-  sebum_counts %>% filter(Extreme_Sebum == "Medium") %>% pull(n)
high_n <- sebum_counts %>% filter(Extreme_Sebum == "High") %>% pull(n)
```

```{r}
# PERMANOVA
dist_metabolites <- vegdist(data_sample_extreme_clr, method = "euclidean")
disper_donor <- betadisper(dist_metabolites, PCA_whole_scores$Subject_ID)

#significant variation between the groups
anova(disper_donor)
permanova_sebumgroup <- adonis2(dist_metabolites ~ PCA_whole_scores$Subject_ID + PCA_whole_scores$Extreme_Sebum, PCA_whole_scores, na.action = na.omit, by="margin")
```

```{r}
sebum_row <- permanova_sebumgroup[2, ]
r2 <- round(sebum_row$R2 * 100, 1)
f_val <- round(sebum_row$F, 2)
p_val <- format(sebum_row$`Pr(>F)`, digits = 3, scientific = TRUE)

label_text <- paste0("PERMANOVA\nR² = ", r2, "%\nF = ", f_val, "\np = ", p_val)

```

```{r}
# Create PCA plot
PCA_plot_sebum_extreme <- PCA_whole_scores %>%
  ggscatter(x = "PC1", y = "PC2", color = i, alpha = 0.6,
            title = "PCA of RCLR-Transformed Metabolite Features by Sebum Level",
            xlab = paste("PC1 (", round(PCA_whole$prop_expl_var$X[1]*100, 1), "%)", sep = ""), 
            ylab = paste("PC2 (", round(PCA_whole$prop_expl_var$X[2]*100, 1), "%)", sep = ""),
            ggtheme = theme_classic()) +
  
  geom_point(data = PCA_whole_scores %>%
               group_by((!!sym(i))) %>%
               summarise(across(matches("PC"), mean)), 
             aes(PC1, PC2, color = (!!sym(i))), size = 4, shape = 8) +
  
  scale_color_manual(
    name = "Sebum Group",
    values = c("Low" = "#4575b4", "High" = "#e08214"),
    labels = c(
      Low = paste0("Low (< 3.5 µg/cm²) (n=", low_n, ")"), #7
      High = paste0("High (> 16.9 µg/cm²) (n=", high_n, ")") #20
    )
  ) +
  
  theme(plot.title = element_text(size = 15), 
        axis.title = element_text(size = 12),
        axis.text = element_text(size = 12),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 11)) +
  
  coord_fixed()+ annotate("text", 
    x = Inf, y = Inf,
    label = label_text,
    hjust = 1.1, vjust = 1.2, size = 3.8
  )
```


```{r}
# PLSDA
PLSDA_extreme <- mixOmics::plsda(data_sample_extreme_clr, PCA_whole_scores$Extreme_Sebum, ncomp = 5, scale = TRUE)
PLSDA_extreme_scores <- data.frame(PLSDA_extreme$variates$X) %>%
  rownames_to_column("SampleID") %>%
  left_join(metadata_with_runorder_clean, by = c("SampleID" = "filename_2"))



# Create PLS-DA plot
PLSDA_extreme_plot <- PLSDA_extreme_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "Extreme_Sebum", alpha = 0.6,
            title = "PLS-DA of RCLR-Transformed Metabolite Features by Sebum Level",
            xlab = paste("Component 1 (", round(PLSDA_extreme$prop_expl_var$X[1] * 100, 1), "%)", sep = ""),
            ylab = paste("Component 2 (", round(PLSDA_extreme$prop_expl_var$X[2] * 100, 1), "%)", sep = ""),
            ggtheme = theme_classic()) +
  
  geom_point(
    data = PLSDA_extreme_scores %>%
      group_by(Extreme_Sebum) %>%
      summarise(across(matches("comp"), mean)),
    aes(comp1, comp2, color = Extreme_Sebum),
    size = 4, shape = 8
  ) +
  
  scale_color_manual(
    name = "Sebum Group",
    values = c("Low" = "#4575b4", "High" = "#e08214"),
    labels = c(
      Low = paste0("Low (< 3.5 µg/cm²) (n=", low_n, ")"),
      High = paste0("High (> 16.9 µg/cm²) (n=", high_n, ")")
    )
  ) +
  
  theme(
    plot.title = element_text(size = 15),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  
  coord_fixed() 
```
```{r}
# Save ggplot object as an RDS file
saveRDS(PLSDA_extreme_plot, file = "~/OneDrive - UGent/CMET/Shiseido_Project/Figures/Metabolomics/PLSDA_extreme_plot.rds")

```


```{r}
Loadings_extreme <- plotLoadings(PLSDA_extreme, contrib = "max") %>%
  as.data.frame(check.names = FALSE) %>%
  rownames_to_column()
```

```{r}
set.seed(123)
# Dynamically remove 'X.' prefix if it exists
colnames(Loadings_extreme) <- gsub("^X\\.", "", colnames(Loadings_extreme))

# Now, select the columns of interest
Loadings_extreme <- Loadings_extreme %>%
  dplyr::select(rowname, GroupContrib, importance)

perf_PLSDA_extreme <- perf(PLSDA_extreme, validation = "Mfold", folds =5, nrepeat=100, auc=TRUE) #perform 100 iunstead of 10 M-fold or loo
plot(perf_PLSDA_extreme, legend = FALSE)
```

```{r}

png("../figures/supplementary/suppl_fig11.png", res = 300, width = 5, height = 5, units = "in")
 
# Generate the PLS-DA performance plot
plot(perf_PLSDA_extreme)

# Close the device to save the plot
dev.off()

```

```{r}
print(perf_PLSDA_extreme$error.rate)
print(perf_PLSDA_extreme$auc)

```
```{r}
colnames(perf_PLSDA_extreme)
```


```{r}
#when excluding chin, ear, neck, temple ; ONLY FOREHEAD, NOSE CHEEK
label_text <- paste0("AUC = 0.92, BER = 0.14")
```


```{r}
# Create PLS-DA plot
PLSDA_extreme_plot <- PLSDA_extreme_scores %>%
  ggscatter(x = "comp1", y = "comp2", color = "Extreme_Sebum", alpha = 0.6,
            title = "PLS-DA of RCLR-Transformed Metabolite Features by Sebum Level",
            xlab = paste("Component 1 (", round(PLSDA_extreme$prop_expl_var$X[1] * 100, 1), "%)", sep = ""),
            ylab = paste("Component 2 (", round(PLSDA_extreme$prop_expl_var$X[2] * 100, 1), "%)", sep = ""),
            ggtheme = theme_classic()) +
  
  geom_point(
    data = PLSDA_extreme_scores %>%
      group_by(Extreme_Sebum) %>%
      summarise(across(matches("comp"), mean)),
    aes(comp1, comp2, color = Extreme_Sebum),
    size = 4, shape = 8
  ) +
  
  scale_color_manual(
    name = "Sebum Group",
    values = c("Low" = "#4575b4", "High" = "#e08214"),
    labels = c(
      Low = paste0("Low (< 3.5 µg/cm²) (n=", low_n, ")"),
      High = paste0("High (> 16.9 µg/cm²) (n=", high_n, ")")
    )
  ) +
  
  theme(
    plot.title = element_text(size = 15),
    axis.title = element_text(size = 12),
    axis.text = element_text(size = 12),
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 11)
  ) +
  
  coord_fixed() +
  
  coord_fixed()+ annotate("text", 
    x = Inf, y = Inf,
    label = label_text,
    hjust = 1.1, vjust = 1.2, size = 3.8
  )
```

```{r}
VIPs_extreme <- as.data.frame(mixOmics::vip(PLSDA_extreme))
VIPs_extreme_filter <- dplyr::filter(VIPs_extreme, VIPs_extreme$comp1 > 1) # larger than 1 is seen as an important component
VIPs_extreme_filter$ID <- rownames(VIPs_extreme_filter)
VIPs_extreme_select <- VIPs_extreme_filter %>% dplyr::select(ID, comp1)
VIPs_extreme_Load <- VIPs_extreme_select %>% 
  left_join(Loadings_extreme, by = c("ID" = "rowname")) %>% arrange(desc(comp1)) %>%
  left_join(info_feature_complete, by = c("ID" = "Feature")) %>% dplyr::select(1:8)
```



```{r}

write_csv(x = VIPs_extreme_Load, file = "../output/metabolomics/PLSDA/VIPs_extreme_Load_data.csv")

```

```{r}
VIPs_extreme_Load <- read_csv("../output/metabolomics/PLSDA/VIPs_extreme_Load_data_SZ_MC_PD_05072025.csv")
```

```{r}
VIPs_extreme_Load2 <- VIPs_extreme_Load %>%
  mutate(
    cytoscape_annotation = if_else(
      is.na(Compound_Name) | Compound_Name == "",
      paste0("(m/z: ", round(mz, 2), ", RT: ", round(RT, 2), ")"),
      paste0(Compound_Name_2, " (m/z: ", round(mz, 2), ", RT: ", round(RT, 2), ")")
    )
  )
```


```{r}
#adapted it for Compound_Name2
VIPs_anno <- VIPs_extreme_Load %>%
  filter(!is.na(Compound_Name_2))

# Set compound names as rownames
rownames(VIPs_anno) <- VIPs_anno$Compound_Name_2


```

```{r}

#selected_ids <- c(2059,2287, 2069, 1921, 45, 2441, 2509, 2245)
selected_ids <- c(2314,1744, 2069, 1883, 2409, 2509, 2287, 2160)
selected_ids <- c(2314,1744, 2069, 1883, 2409, 2509, 2287, 2160,1536,1011) #to remove diethylphtalate and maybe also diphenhydramine
VIPs_selected <- VIPs_anno %>% filter(ID %in% selected_ids)
VIPs_remaining <- VIPs_anno %>% filter(!(ID %in% selected_ids))

```
```{r}
# Recreate sebum_group thresholds (same as earlier in your code)
q          <- quantile(metadata_with_runorder_clean$sebum, probs = c(0.25, 0.75), na.rm = TRUE)
low_thresh  <- q[1]
high_thresh <- q[2]

# Build clr_long with sebum_group created inline
clr_long <- data_sample_clr %>%
  as.data.frame() %>%
  rownames_to_column("SampleID") %>%
  left_join(
    metadata_with_runorder_clean %>%
      dplyr::select(filename_2, sebum) %>%
      mutate(
        sebum_group = case_when(
          sebum <  low_thresh  ~ "low",
          sebum <= high_thresh ~ "medium",
          sebum >  high_thresh ~ "high"
        )
      ),
    by = c("SampleID" = "filename_2")
  ) %>%
  rename(group = sebum_group) %>%
  filter(!is.na(group))

# Counts per group
sebum_counts <- clr_long %>% count(group)
low_n    <- sebum_counts %>% filter(group == "low")    %>% pull(n)
medium_n <- sebum_counts %>% filter(group == "medium") %>% pull(n)
high_n   <- sebum_counts %>% filter(group == "high")   %>% pull(n)

colors <- c("low" = "#4575b4", "medium" = "#fee0b6", "high" = "#e08214")
```



```{r}
make_vip_boxplots <- function(vip_df, clr_data, output_prefix) {
  plots <- list()

  for (i in seq_len(nrow(vip_df))) {
    metabolite     <- as.character(vip_df$ID[i])
    metabolite_name <- vip_df$Compound_Name_2[i]
    mz_value       <- round(vip_df$mz[i], 2)
    rt_value       <- round(vip_df$RT[i], 2)

    clr_data$group <- factor(clr_data$group, levels = c("low", "medium", "high"))

    temp_data <- clr_data %>%
      dplyr::select(group, !!sym(metabolite)) %>%
      filter(!is.na(.data[[metabolite]]))

    kw_res  <- kruskal.test(temp_data[[metabolite]] ~ temp_data$group)
    p_val   <- kw_res$p.value
    stars   <- if (p_val < 0.001) "***" else if (p_val < 0.01) "**" else if (p_val < 0.05) "*" else "ns"
    label_text <- paste0("Kruskal-Wallis p = ", signif(p_val, 3), " ", stars)
    y_max   <- max(temp_data[[metabolite]], na.rm = TRUE)

    p <- clr_data %>%
      ggplot(aes(x = group, y = .data[[metabolite]], fill = group, color = group)) +
      geom_boxplot(outlier.shape = NA, color = "black", width = 0.5) +
      geom_jitter(width = 0.2, alpha = 0.6, size = 0.5) +
      scale_fill_manual(values  = colors) +
      scale_color_manual(values = colors) +
      scale_x_discrete(labels = c(
        "low"    = paste0("Low Sebum\n(< 3.5 µg/cm²)\n(n = ",       low_n,    ")"),
        "medium" = paste0("Medium Sebum\n(3.5–16.9 µg/cm²)\n(n = ", medium_n, ")"),
        "high"   = paste0("High Sebum\n(> 16.9 µg/cm²)\n(n = ",     high_n,   ")")
      )) +
      labs(
        y     = paste0("rclr ", metabolite_name, "\n(m/z: ", mz_value, ", RT: ", rt_value, ")"),
        x     = NULL,
        title = metabolite_name
      ) +
      theme_minimal() +
      theme(
        axis.text.x  = element_text(size = 12),
        axis.text.y  = element_text(size = 12),
        axis.title.y = element_text(size = 12),
        plot.title   = element_text(size = 15),
        axis.ticks   = element_line(),
        panel.grid   = element_blank(),
        legend.position = "none",
        axis.line    = element_line(color = "black")
      ) +
      annotate("text", x = 2, y = y_max * 1.1, label = label_text, size = 4)

    plots[[length(plots) + 1]] <- p
  }
  return(plots)
}
```


```{r}
vip_plots <-make_vip_boxplots(VIPs_selected, clr_long, "selected_VIPs")
vip_remaining_plots <-make_vip_boxplots(VIPs_remaining, clr_long, "remaining_VIPs")

```

```{r}
# Select and arrange your desired plots
final_figure <- ggarrange(
  vip_plots[[1]], vip_plots[[4]], vip_plots[[3]],vip_plots[[8]], 
  NULL, NULL, NULL, NULL,
  vip_plots[[2]], vip_plots[[5]], vip_plots[[6]], vip_plots[[7]],
  NULL, NULL, NULL, NULL,
  ncol = 4, nrow = 4,
  labels = c("d", "e", "f", "g", "", "", "", "", "h", "i","j", "k", "", "", "","" ),
  heights = c(1, 0.75, 1, 0.75),
  font.label = list(size = 16, face = "bold")
)

# Show it in the viewer
print(final_figure)
```